---
title: "Metier_efforts"
author: "Emma Fuller"
date: "April 19, 2014"
output:
  ioslides_presentation:
    fig_width: 7.5
    smaller: yes
    transition: faster
  beamer_presentation:
    fonttheme: structuresmallcapsserif
  revealjs_presentation:
    theme: sky
    transition: linear
---
##goal: MCA on entire logbook contents, CLARA on entire dataset

```{r, echo=FALSE,message=FALSE}
require(ca)
require(cluster)
require(ggplot2)
require(plotrix)
setwd("/Volumes/NOAA_Data/CNH/")

# load data
load("Analysis/Metiers/results/2014-03-22/ftl_clust29.Rdata")
```

relevant effort variables to cluster from fish ticket dataset

+ `month`
+ `grid` (finer than `grgroup`)
+ `catch profile`

unsure what `pargrp` is, from [PacFin](http://pacfin.psmfc.org/pacfin_pub/table_cols.php)
> Participation group; see par-group in table cl.]; vesseltype [Vessel type (see vid-type in table cl)])

## Subset data to these variables
```{r}
mca.data <- subset(ftl_clust, select=c("month","grid","catch_profile"))
```

however, issue of unspecified gears:

+ `OTH` = other known gear
+ `OHL` = other hook and line gear
+ `ONT` = other net gear
+ `OPT` = other pot gear
+ `USP` = unknown or unspecified
+ `ODG` = other dredge gear

Clear that I can remove `USP`, but what about others? 

## Distribution of gear types
```{r, echo=FALSE, fig.align='center'}
ggplot(ftl_clust,aes(x=grid)) + geom_bar()
```
Mostly made up of `CPT` (crab pots), `LGL` (long line/set line), `MDT` (midwater trawls), `RLT` (roller trawls)

## Big effect of removals?
How many trips are in each of these "other" categories?

```{r eval=FALSE}
OTH = length(which(ftl_clust$grid=="OTH"))
OHL = length(which(ftl_clust$grid=="OHL"))
ONT = length(which(ftl_clust$grid=="ONT"))
OPT = length(which(ftl_clust$grid=="OPT"))
USP = length(which(ftl_clust$grid=="USP"))


  
mca.analysis <- mjca(mca.data)
plot(mca.analysis,contrib="absolute",mass=c(TRUE,TRUE))
```

