---
title: "Comparing types of diversity"
author: "Emma Fuller"
date: "January 25, 2015"
output: html_document
---
**need to spend some more time with how effective connectance is calculated**


See `comparing_diversities.Rmd` for methods, this is trying to understand the similarities and differences between using the Simpsons index calculated by averaging a port's vessels and using the mean of a port's yearly effective connectance participation plot.

Interested in understanding why the vessel-level diversity plateaus but effective connectance generates differences (ports to the left of the red line below) and those in which efefctive connectance is about the same, but there's a large variance in vessel-level (between yellow lines below).

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(scales)
port_metrics <- read.csv("/Users/efuller/1/CNH/Analysis/Metiers/results/2015-01-18/port_metrics.csv")
port_metrics$X <- NULL

with(port_metrics, scatter.smooth(CZ_95, vessel_level,pch=19,col=alpha("lightgrey",.75),cex=2,bty="n",lwd=3))
with(port_metrics, text(CZ_95, vessel_level, tolower(pcid), cex=.5))
abline(v=9.8, col="tomato",lwd=3, lty=2)
abline(v=2.5, col="goldenrod",lwd=3, lty=2)
abline(v=4.7, col="goldenrod",lwd=3, lty=2)
```

### High `CZ`, intermediate vessel level
```{r,echo=FALSE, warning=FALSE,message=FALSE}
high_CZ <- subset(port_metrics, CZ_all > 10)
tickets <- readRDS("/Users/efuller/1/CNH/Analysis/Metiers/results/2015-01-09/code/3_exploreBuildwebs/tickets.RDS")
source("/Users/efuller/1/CNH/Analysis/Metiers/results/2015-01-25/functions/defineParticipationNetwork.R")
library(igraph);library(reshape2); library(RColorBrewer)
orf = define_participationPlot(port="ORF",graph=TRUE)
mro = define_participationPlot(port="MRO",graph=TRUE)
avl = define_participationPlot(port="AVL",graph=TRUE)
sp = define_participationPlot(port="SP",graph=TRUE)
oxn = define_participationPlot(port="OXN",graph=TRUE)
sb = define_participationPlot(port="SB",graph=TRUE)
osd = define_participationPlot(port="OSD",graph=TRUE)
pcc = define_participationPlot(port="PCC",graph=TRUE)
osl = define_participationPlot(port="OSL",graph=TRUE, restrict=FALSE)

by_year <- read.csv("/Users/efuller/1/CNH/Analysis/Metiers/results/2015-01-18/by_year_port_metrics.csv")

````