---
title: "Fisheries food web"
author: "Emma Fuller"
date: "September 23, 2014"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

```{r load_data, echo=FALSE, message=FALSE, warning=FALSE}
ftl <- read.csv("/Users/efuller/1/CNH/Data/Catch/FTL_2009-2013_w-o-c_samhouri.csv", 
                stringsAsFactors = F)
```

## Introduction
When we manage ecosystems, we manage ourselves, not the organisms that make up the natural resources we covet.[^1] When fishermen are considered, most analyses are focused on understanding and managing the relationship between individual fishermen and a government authority, despite the demonstrated understanding that commmunities are important to the success of whatever management approach is proposed.[^2] 

Communities also receive renewed attention in the Magnuson-Stevens Act which requires that fishing communities are considered before management is changed. And with the shift to focusing on ecosystem services and social-ecological systems, understanding the linkages between individuals, groups and the environment are required for understanding the system holistically. 

Here I present analyses that map out connections among individuals, communities and fish species. Admittedly I take the easiest step of linking together harvesters with the ecological species they target. But my hope is that this can be built upon for a more complete picture of fisheries systems. 

This work is helpful in that it also considers how groups of fishermen may respond to perturbations of the fishery system and the relative amounts of risk each fishermen is exposed to via revenues.  

[^1]: we manage fishermen not fish (Jentoft 1997), Fulton et al. 2011
[^2]: Papers: Hilborn on leadership; Acheson, Wilson on communication being pre-requisite for bottom up success; Jentoft on the importance of community (2000); Pinsky & Fogarty on adaptation to climate

## Methods

The goal of this work is to build a fisheries foodweb for the US West coast. In this web, _predator nodes_ are vessels and _prey nodes_ are fish targets. Here fish targets are defined as assemblages of species commonly caught together. Often these fish targets are single species (i.e. dungeness crab, albacore) but often multiple species are caught (i.e. dover sole, thornyheads, sablefish). Predator nodes are defined as the least common demominator in the existing combinations of gear and fishing targets. 

### Fish targets
I use fish tickets to define fish targets by looking for clusters of catches with similar catch profiles. Preliminary work [**Appendix**] suggests that gear constrains possible catch.[^3] For computational tractability, I subdivide the fish tickets by gear group[^4] and search for patterns in catch composition (see table below).

[^3]: Alternatively, this is what metiers do.
[^4]: See [PacFin](http://pacfin.psmfc.org/pacfin_pub/data_rpts_pub/code_lists/gr_tree.txt)

Gear grouping | gears included
------------- | --------------
Mobile trawl  | roller trawls, flatfish trawls, groundfish trawls
Fixed gear    | long-lines, fish pots
Troll         | troll
Crab pots     | crab pots
Shrimp trawl  | double rigged, single rigged
Midwater trawl | midwater trawl
**FIX ME**    | **put in real groups!** 

```{r split_grgroup, echo=FALSE, message=FALSE, warning=FALSE}
source("/Users/efuller/1/CNH/Analysis/Metiers/results/2014-09-23/helper_funs.R")
par(mfrow=c(1,3))
output <- find_target("TWL",2012)

V(output[["cg2"]])$name <- c("arrow_dover","whiting","bait_shrimp","cal_hal","cukes","chinook","chili_petrale_plus","rockfish_risky","lingcod_other_rockfish", "osrm","crab","pac_halibut","ridge_prawn", "canary_nusf","scrp.horn.turbot","sable","coho")

plot_targets(output[["cg2"]])

out_bip <- bipartite_network(port_trips = output[["port_trips"]], bc_mat = output[["bc_mat"]], ic = output[["ic"]], cg2 = output[["cg2"]], scale_edge = 800)

participation_network(trips = out_bip[["trips"]], trip_gear = out_bip[["trip_gear"]])
```

## Grouping vessels by yearly catch
Here I subset by year, and aggregate catch for each vessel. I then calculate a pairwise bray-curtis dissimilarity metric and use that to build a weighted, undirected network. Using this network I apply InfoMap to look for modules. These modules are considered "fishing strategies": groupds of vessels that catch similar types of species over the course of a year. 

Using these strategies, I can calculate statistics on the variance and mean yearly revenue. 

Here I subset to consider only strategies in which at least 5 vessels participated. This is partially for confidentiality reasons (I can't show figures of data that describe fewer than 3 vessels), and partially because we want to focus on the major fisheries. However I realize the a strategy having few vessels is different than a fishery having few participants. These strategies may be uncommon combinations of fisheries that are quite large. This needs more thought. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(plyr)
source("/Users/efuller/1/CNH/Analysis/Metiers/results/2014-09-23/helper_funs.R")
slopes <- list()
strategies <- list()
years <- 2009:2013
par(mfrow=c(2,3))
for(i in 1:length(years)){
  #cat(years[i],":\n")
  out_strategies <- find_strategies(year = years[i])
  #cat("strategies found, ")
  strategies[[i]] <- calculate_strategy_stats(ic = out_strategies[["ic"]], 
                                       port_trips = out_strategies[["port_trips"]], 
                                       cast_trips=out_strategies[["cast_trips"]])
  #cat("stats calculated, ")
  plot_strategies(strategies[[i]][["strategy"]], years[i], yr_trips = strategies[[i]][["yr_trips"]])
  slopes[[i]] <- get_slope(strategy = strategies[[i]][["strategy"]])
  #cat("slope found. done.\n")

}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#library(ggvis)
#df <- data.frame(year = factor(years), slope = unlist(slopes))
#df %>% ggvis(~year, ~slope) %>% layer_lines() %>% layer_points()

plot(years, unlist(slopes), type = "o",pch=19, bty="n", xlab="Year",ylab="slope", asp = 0)
```

Next is to plot diversity against $\frac{\log{\text{mean revenue}}}{\log{\text{sd revenue}}}$. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(2,3))
for(i in 1:length(strategies)){
  plot(log(strategies[[i]][["strategy"]]$mean_rev)/log(strategies[[i]][["strategy"]]$sd_rev), strategies[[i]][["strategy"]]$mean_div, type="p",pch = 19, bty = "n", xlab = "log(mean revenue)/log(sd revenue)", ylab = "mean sw index", main = years[i])
}
```
The take away here is that maybe the more diverse your strategy (in terms of species) the better you do. Maybe.

### Redoing it with metiers
Rather than aggregating strictly on species, looking at the number of metiers vessels participate within. This approach will correct for the fact that trawling may make a vessel's yearly catch species rich, when in fact they are still participating in one "fishery" as commonly defined. 

Each year split fish tickets into gear groups and look for ecological targets. Then look at the combination of ecological targets a vessel focused on during a year and repeat the same analysis as above. 

To do this I need to make sure that only one gear is associated with one vessel. However there are about 3000 trips which list more than one gear. For now I will drop those trips until I can figure out something better to do with them. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# find the vessels which have more than 1 gear listed
library(dplyr)
just_gear <- select(ftl, ftid, grid)
just_gear <- unique(just_gear)
gear_types <- table(just_gear$ftid, just_gear$grid)
total_gear <- rowSums(gear_types)
extra_gear_trips <- names(total_gear)[which(total_gear > 1)]

ftl_trips <- subset(ftl, !(ftid %in% extra_gear_trips))

#find metiers by year and by gear

gear_groups <- unique(ftl_trips$grgroup)
years <- unique(ftl_trips$year)
metier_trips <- list()
for(i in 1:length(gear_groups)){
  cat(gear_groups[i],":\n")
  for(j in 1:length(years)){
    cat(years[j],":\n")
    metier_trips[[i]] <- find_metiers(tickets = ftl_trips, year =years[j], 
                                 gear_group = gear_groups[i], message = "YES")
    }
    cat("\n")
  }
}

library(data.table)
ftl_dt

```

```{r}
# group strategies


```



### Cost distributions
A more diverse set of species should reduce risk only if their distributions overlap. If species price distributions do not overlap, variation will increase with diversity. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
sub_ftl  <- subset(ftl, spid %in% names(sort(table(ftl$spid), decreasing=T)[which(sort(table(ftl$spid), decreasing=T)>2850)]))
ggplot(sub_ftl, aes(factor(spid), ppp)) + geom_violin(aes(fill = factor(spid)), scale = "width") + scale_y_continuous(limits = c(0, 35)) +guides(fill=FALSE) + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +xlab("species") + ylab("$/lb")
```

## Appendix
### A1: Relevant gear groupings
Because the dataset is so large (> 500,000 trips), I consider subsets of trips for computational tractability. The three ports most which had the most landings were Santa Barbara, CA (SB); Newport, OR (NEW); Westport, WA (WPT); and Oxnard, CA (OXN). 



```{r plot_port_count, echo=FALSE, message=FALSE,warning=FALSE}
by_port <- ftl[!duplicated(ftl$ftid),]
count_port <- sort(table(by_port$pcid),decreasing=T)
paint <- rep("grey",90)
paint[1:4] <- "tomato"
barplot(count_port, bor=F, las = 2,cex.names = .25, col=paint, ylab = "Trips landed: 2009-2013", xlab = "Port",cex.axis = .75)
legend("topright", fill=c("tomato"), legend = c("ports examined"), bty="n", border="tomato")
```

### Santa Barbara

```{r target_infomap_SB, echo=FALSE, message=FALSE, warning=FALSE}
library(igraph); library(vegan)

port_trips <- subset(ftl, pcid=="SB" & year == 2012, select = c("veid","ftid","spid","ppp","landed_wt","tdate","grid"))
  library(vegan); library(reshape2)
  melt_trips <- melt(port_trips, id.vars = c("veid","ftid","spid","tdate","grid"), measure.vars = "landed_wt")
  cast_trips <- dcast(melt_trips, ftid ~ spid, fun.aggregate = sum)
  rownames(cast_trips) <- cast_trips$ftid
  cast_trips$ftid <- NULL
  bc <- vegdist(cast_trips, method = "bray")
  bc_sim <- abs(bc-1)
  bc_mat <- as.matrix(bc_sim)
  
  library(igraph)
  bc_net <- graph.adjacency(bc_mat, mode="undirected", weighted = TRUE)
  ic <- infomap.community(bc_net, e.weights = E(bc_net)$weight)

  cg <- contract.vertices(bc_net, membership(ic))
  #E(cg)$weight <- 1 # leave this off, let the weights add up as they are. 
  cg2 <- simplify(cg, remove.loops=TRUE, edge.attr.comb = median)
```

Subsetting to Santa Barbara for 2012 I'm left with `r length(unique(port_trips$ftid))` trips. Using the fish tickets I calculate a pair-wise Bray-Curtis index of dissimilarity between all trips. After converting this to a similarity, I build an undirected, weighted network in which the edge weights are the similarity between trips. With this network I use InfoMap to search for communities, or modules. 

The network below is produced by aggregating trips into communities. Each node represents a characteristic fishing target, and is informally labeled by the dominant species present. The edge width is the median similarity between nodes $i$ and node $j$. Thus it is a rough proxy for the overlap in species composition between nodes. 

```{r SB_target_plot, echo=FALSE, warning=FALSE, message=FALSE}
V(cg2)$name <- c("red_urchin","rock_crab","spiny_lobster_rock_crab","sea_cukes","chili_etal","sable","seabass", "cabezon_grass","vermillion_etal","ridge_prawn","chinook","spot_prawn","hagfish","swordfish","albacore","other_crab","other_urchins","crab","other_mollusk","yellowtail_dolphin","other_skate","squid", "starfish")

V(cg2)$size <- table(membership(ic))
par(bg="steelblue")
library(scales)
l <- layout.fruchterman.reingold(cg2, niter=1000, area=vcount(cg2)^3, repulserad=vcount(cg2)^3)
par(mai = rep(0,4), bg="white")
plot(cg2, vertex.size = log(V(cg2)$size)*2.5, layout = l, edge.width = E(cg2)$weight*20, edge.label = "", vertex.label.font=1, vertex.label.color="black", vertex.frame.color="steelblue", vertex.color = "steelblue", edge.color=alpha("grey",1-E(cg2)$weight), vertex.label.family = "sans", vertex.label.cex = .75)

```

When these targets are connected to gears we find that there is little overlap among gears, i.e. few gears share the same target. This is despite the fact that there is overlap among catches in species composition. This suggests that gear broadly determines the possible fishing targets.

```{r bibartite_graph_SB, echo=FALSE, message=FALSE, warning=FALSE}
# assign target cluster to each trip
trips <- port_trips
trips$target <- rep(NA, nrow(trips))
for(i in 1:length(communities(ic))){
  trips$target[which(trips$ftid %in% rownames(bc_mat[communities(ic)[[i]],]))] <- V(cg2)$name[[i]]
}

# now make bipartite network for gear
trip_types <- table(trips$target, trips$grid)
trip_gear <- melt(trip_types)
trip_gear <- subset(trip_gear, value > 10)
colnames(trip_gear) <- c("from","to","weight")

fisheries <- graph.data.frame(trip_gear, directed = FALSE)
E(fisheries)$weight <- trip_gear$weight
num_targets <- length(unique(trip_gear$from))
num_gears <- length(unique(trip_gear$to))
V(fisheries)$type <- c(rep(TRUE,num_targets), rep(FALSE, num_gears))
V(fisheries)$color <- c(rep("steelblue", num_targets), rep("brown", num_gears))
par(mai=rep(0,4), bg="white")
plot(fisheries, edge.width = E(fisheries)$weight/100, layout=layout.bipartite, ylim=c(-1,1),vertex.label.cex=.5, vertex.label.color="white", vertex.label.family = "sans", vertex.frame.color=V(fisheries)$color)
```

Looking at which vessels participate in which fisheries I find

```{r participation_graph_SP, echo=FALSE, message=FALSE, warning=FALSE}
# now make participation network. 
trips$fishery <- paste(trips$target, trips$grid, sep="_")
# restrict to fisheries used
used_fisherys <- unique(paste(trip_gear$from, trip_gear$to, sep="_"))
used_trips <- subset(trips, fishery %in% used_fisherys)
by_fishery <- table(used_trips$veid, used_trips$fishery)

fishery_mat <- matrix(ncol=length(used_fisherys), nrow=length(used_fisherys))
colnames(fishery_mat) <- colnames(by_fishery)
rownames(fishery_mat) <- colnames(by_fishery)

for(i in 1:ncol(by_fishery)){ # for each fishery except the last one
  fish_i <- colnames(by_fishery)[i]
  #cat("j: \n")
  for(j in i:ncol(by_fishery)){
    fish_j <- colnames(by_fishery)[j]
    num_ves = length(which(which(by_fishery[,i]>0) %in% which(by_fishery[,j] > 0)))
    
    mat_row <- which(rownames(fishery_mat)==fish_j)
    mat_col <- which(colnames(fishery_mat)==fish_i)
    
    fishery_mat[mat_row, mat_col] <- num_ves
    #cat(j," ")
  }
  #cat("\n")
}

g<-graph.adjacency(fishery_mat,weighted=T, diag=F, mode = "lower")
l <- layout.fruchterman.reingold(g, niter=500, area=vcount(g)^2, repulserad=vcount(g)^2)
#plot(g, edge.width = E(g)$weight, layout = l)

# looking for communities of vessels 

ic_gear <- infomap.community(g)
library(RColorBrewer)
paint <- brewer.pal(6,"Dark2")
V(g)$membership <- ic_gear$membership
V(g)[membership==1]$color <- paint[1]
V(g)[membership==1]$frame.color <- paint[1]
V(g)[membership==2]$color <-paint[2]
V(g)[membership==2]$frame.color <-paint[2]
V(g)[membership==3]$color <- paint[3]
V(g)[membership==3]$frame.color <-paint[3]
V(g)[membership==4]$color <-paint[4]
V(g)[membership==4]$frame.color <-paint[4]
V(g)[membership==5]$color <- paint[5]
V(g)[membership==5]$frame.color <- paint[5]
V(g)[membership==6]$color <- paint[6]
V(g)[membership==6]$frame.color <- paint[6]
par(mai=rep(0,4))

V(g)$size <- colSums(by_fishery)

plot(g, vertex.color=V(g)$color, edge.width = E(g)$weight, vertex.size=log(V(g)$size)*3, vertex.frame.color=V(g)$frame.color, vertex.label.color="black",layout=l, vertex.label.family = "sans", vertex.label.cex = .75)
```

### Westport

```{r target_infomap_WPT, echo=FALSE, message=FALSE, warning=FALSE}
library(igraph); library(vegan)

port_trips <- subset(ftl, pcid=="WPT" & year == 2012, select = c("veid","ftid","spid","ppp","landed_wt","tdate","grid"))
  melt_trips <- melt(port_trips, id.vars = c("veid","ftid","spid","tdate","grid"), measure.vars = "landed_wt")
  cast_trips <- dcast(melt_trips, ftid ~ spid, fun.aggregate = sum)
  rownames(cast_trips) <- cast_trips$ftid
  cast_trips$ftid <- NULL
  bc <- vegdist(cast_trips, method = "bray")
  bc_sim <- abs(bc-1)
  bc_mat <- as.matrix(bc_sim)
  
  library(igraph)
  bc_net <- graph.adjacency(bc_mat, mode="undirected", weighted = TRUE)
  ic <- infomap.community(bc_net, e.weights = E(bc_net)$weight)

  cg <- contract.vertices(bc_net, membership(ic))
  #E(cg)$weight <- 1 # leave this off, let the weights add up as they are. 
  cg2 <- simplify(cg, remove.loops=TRUE, edge.attr.comb = median)
```


```{r WPT_target_plot, echo=FALSE, warning=FALSE, message=FALSE}
V(cg2)$name <- c("crab","albc_nanc","pelagics","chinook","pink_shrimp","razor_clam","whiting", "sable","anchovy","spotted_prawn","hagfish","oyster")

V(cg2)$size <- table(membership(ic))
l <- layout.fruchterman.reingold(cg2, niter=1000, area=vcount(cg2)^3, repulserad=vcount(cg2)^3)
par(mai = rep(0,4), bg="white")
plot(cg2, vertex.size = log(V(cg2)$size)*2.5, layout = l, edge.width = E(cg2)$weight*20, edge.label = "", vertex.label.font=1, vertex.label.color="black", vertex.frame.color="steelblue", vertex.color = "steelblue", edge.color=alpha("grey",1-E(cg2)$weight), vertex.label.family = "sans", vertex.label.cex = .75)

```

```{r bibartite_graph_WPT, echo=FALSE, message=FALSE, warning=FALSE}
# assign target cluster to each trip
trips <- port_trips
trips$target <- rep(NA, nrow(trips))
for(i in 1:length(communities(ic))){
  trips$target[which(trips$ftid %in% rownames(bc_mat[communities(ic)[[i]],]))] <- V(cg2)$name[[i]]
}

# now make bipartite network for gear
trip_types <- table(trips$target, trips$grid)
trip_gear <- melt(trip_types)
trip_gear <- subset(trip_gear, value > 10)
colnames(trip_gear) <- c("from","to","weight")

fisheries <- graph.data.frame(trip_gear, directed = FALSE)
E(fisheries)$weight <- trip_gear$weight
num_targets <- length(unique(trip_gear$from))
num_gears <- length(unique(trip_gear$to))
V(fisheries)$type <- c(rep(TRUE,num_targets), rep(FALSE, num_gears))
V(fisheries)$color <- c(rep("steelblue", num_targets), rep("brown", num_gears))
par(mai=rep(0,4), bg="white")
plot(fisheries, edge.width = E(fisheries)$weight/100, layout=layout.bipartite, ylim=c(-1,1),vertex.label.cex=.75, vertex.label.color="white", vertex.label.family = "sans", vertex.frame.color=V(fisheries)$color)
```

```{r participation_graph_WPT, echo=FALSE, message=FALSE, warning=FALSE}
# now make participation network. 
trips$fishery <- paste(trips$target, trips$grid, sep="_")
# restrict to fisheries used
used_fisherys <- unique(paste(trip_gear$from, trip_gear$to, sep="_"))
used_trips <- subset(trips, fishery %in% used_fisherys)
by_fishery <- table(used_trips$veid, used_trips$fishery)

fishery_mat <- matrix(ncol=length(used_fisherys), nrow=length(used_fisherys))
colnames(fishery_mat) <- colnames(by_fishery)
rownames(fishery_mat) <- colnames(by_fishery)

for(i in 1:ncol(by_fishery)){ # for each fishery except the last one
  fish_i <- colnames(by_fishery)[i]
  #cat("j: \n")
  for(j in i:ncol(by_fishery)){
    fish_j <- colnames(by_fishery)[j]
    num_ves = length(which(which(by_fishery[,i]>0) %in% which(by_fishery[,j] > 0)))
    
    mat_row <- which(rownames(fishery_mat)==fish_j)
    mat_col <- which(colnames(fishery_mat)==fish_i)
    
    fishery_mat[mat_row, mat_col] <- num_ves
    #cat(j," ")
  }
  #cat("\n")
}

g<-graph.adjacency(fishery_mat,weighted=T, diag=F, mode = "lower")
l <- layout.fruchterman.reingold(g, niter=500, area=vcount(g)^2, repulserad=vcount(g)^2)
#plot(g, edge.width = E(g)$weight, layout = l)

# looking for communities of vessels 

ic_gear <- infomap.community(g)
library(RColorBrewer)
paint <- brewer.pal(6,"Dark2")
V(g)$membership <- ic_gear$membership
V(g)[membership==1]$color <- paint[1]
V(g)[membership==1]$frame.color <- paint[1]
V(g)[membership==2]$color <-paint[2]
V(g)[membership==2]$frame.color <-paint[2]
V(g)[membership==3]$color <- paint[3]
V(g)[membership==3]$frame.color <-paint[3]
V(g)[membership==4]$color <-paint[4]
V(g)[membership==4]$frame.color <-paint[4]
V(g)[membership==5]$color <- paint[5]
V(g)[membership==5]$frame.color <- paint[5]
V(g)[membership==6]$color <- paint[6]
V(g)[membership==6]$frame.color <- paint[6]
par(mai=rep(0,4))

V(g)$size <- colSums(by_fishery)

plot(g, vertex.color=V(g)$color, edge.width = E(g)$weight, vertex.size=log(V(g)$size)*3, vertex.frame.color=V(g)$frame.color, vertex.label.color="black",layout=l, vertex.label.family = "sans", vertex.label.cex = .75)
```

### Newport

```{r target_infomap_NEW, echo=FALSE, message=FALSE, warning=FALSE}
library(igraph); library(vegan)

port_trips <- subset(ftl, pcid=="NEW" & year == 2012, select = c("veid","ftid","spid","ppp","landed_wt","tdate","grid"))
  melt_trips <- melt(port_trips, id.vars = c("veid","ftid","spid","tdate","grid"), measure.vars = "landed_wt")
  cast_trips <- dcast(melt_trips, ftid ~ spid, fun.aggregate = sum)
  rownames(cast_trips) <- cast_trips$ftid
  cast_trips$ftid <- NULL
  bc <- vegdist(cast_trips, method = "bray")
  bc_sim <- abs(bc-1)
  bc_mat <- as.matrix(bc_sim)
  
  bc_net <- graph.adjacency(bc_mat, mode="undirected", weighted = TRUE)
  ic <- infomap.community(bc_net, e.weights = E(bc_net)$weight)

  cg <- contract.vertices(bc_net, membership(ic))
  #E(cg)$weight <- 1 # leave this off, let the weights add up as they are. 
  cg2 <- simplify(cg, remove.loops=TRUE, edge.attr.comb = median)
```

```{r NEW_target_plot, echo=FALSE, warning=FALSE, message=FALSE}
V(cg2)$name <- c("chinook","crab","albacore","sable","whiting","pink_shrimp","lingcod", "hagfish","halibut","octopus","small_halibut","other_crab", "misc_fish","nusp","yellowtail","rockcrab","red_urchin","cockle")

V(cg2)$size <- table(membership(ic))
l <- layout.fruchterman.reingold(cg2, niter=1000, area=vcount(cg2)^3, repulserad=vcount(cg2)^3)
par(mai = rep(0,4), bg="white")
plot(cg2, vertex.size = log(V(cg2)$size)*2.5, layout = l, edge.width = E(cg2)$weight*20, edge.label = "", vertex.label.font=1, vertex.label.color="black", vertex.frame.color="steelblue", vertex.color = "steelblue", edge.color=alpha("grey",1-E(cg2)$weight), vertex.label.family = "sans", vertex.label.cex = .75)

```

```{r bibartite_graph_NEW, echo=FALSE, message=FALSE, warning=FALSE}
# assign target cluster to each trip
trips <- port_trips
trips$target <- rep(NA, nrow(trips))
for(i in 1:length(communities(ic))){
  trips$target[which(trips$ftid %in% rownames(bc_mat[communities(ic)[[i]],]))] <- V(cg2)$name[[i]]
}

# now make bipartite network for gear
trip_types <- table(trips$target, trips$grid)
trip_gear <- melt(trip_types)
trip_gear <- subset(trip_gear, value > 10)
colnames(trip_gear) <- c("from","to","weight")

fisheries <- graph.data.frame(trip_gear, directed = FALSE)
E(fisheries)$weight <- trip_gear$weight
num_targets <- length(unique(trip_gear$from))
num_gears <- length(unique(trip_gear$to))
V(fisheries)$type <- c(rep(TRUE,num_targets), rep(FALSE, num_gears))
V(fisheries)$color <- c(rep("steelblue", num_targets), rep("brown", num_gears))
par(mai=rep(0,4), bg="white")
plot(fisheries, edge.width = E(fisheries)$weight/300, layout=layout.bipartite, ylim=c(-1,1),vertex.label.cex=.5, vertex.label.color="white", vertex.label.family = "sans", vertex.frame.color=V(fisheries)$color)
```

```{r participation_graph_NEW, echo=FALSE, message=FALSE, warning=FALSE}
# now make participation network. 
trips$fishery <- paste(trips$target, trips$grid, sep="_")
# restrict to fisheries used
used_fisherys <- unique(paste(trip_gear$from, trip_gear$to, sep="_"))
used_trips <- subset(trips, fishery %in% used_fisherys)
by_fishery <- table(used_trips$veid, used_trips$fishery)

fishery_mat <- matrix(ncol=length(used_fisherys), nrow=length(used_fisherys))
colnames(fishery_mat) <- colnames(by_fishery)
rownames(fishery_mat) <- colnames(by_fishery)

for(i in 1:ncol(by_fishery)){ # for each fishery except the last one
  fish_i <- colnames(by_fishery)[i]
  #cat("j: \n")
  for(j in i:ncol(by_fishery)){
    fish_j <- colnames(by_fishery)[j]
    num_ves = length(which(which(by_fishery[,i]>0) %in% which(by_fishery[,j] > 0)))
    
    mat_row <- which(rownames(fishery_mat)==fish_j)
    mat_col <- which(colnames(fishery_mat)==fish_i)
    
    fishery_mat[mat_row, mat_col] <- num_ves
    #cat(j," ")
  }
  #cat("\n")
}

g<-graph.adjacency(fishery_mat,weighted=T, diag=F, mode = "lower")
l <- layout.fruchterman.reingold(g, niter=500, area=vcount(g)^2, repulserad=vcount(g)^2)
#plot(g, edge.width = E(g)$weight, layout = l)

# looking for communities of vessels 

ic_gear <- infomap.community(g)
library(RColorBrewer)
paint <- brewer.pal(6,"Dark2")
V(g)$membership <- ic_gear$membership
V(g)[membership==1]$color <- paint[1]
V(g)[membership==1]$frame.color <- paint[1]
V(g)[membership==2]$color <-paint[2]
V(g)[membership==2]$frame.color <-paint[2]
V(g)[membership==3]$color <- paint[3]
V(g)[membership==3]$frame.color <-paint[3]
V(g)[membership==4]$color <-paint[4]
V(g)[membership==4]$frame.color <-paint[4]
V(g)[membership==5]$color <- paint[5]
V(g)[membership==5]$frame.color <- paint[5]
V(g)[membership==6]$color <- paint[6]
V(g)[membership==6]$frame.color <- paint[6]
par(mai=rep(0,4))

V(g)$size <- colSums(by_fishery)

plot(g, vertex.color=V(g)$color, edge.width = E(g)$weight, vertex.size=log(V(g)$size)*3, vertex.frame.color=V(g)$frame.color, vertex.label.color="black",layout=l, vertex.label.family = "sans", vertex.label.cex = .75)
```

### Oxnard
```{r target_infomap_OXN, echo=FALSE, message=FALSE, warning=FALSE}
library(igraph); library(vegan)

port_trips <- subset(ftl, pcid=="OXN" & year == 2012, select = c("veid","ftid","spid","ppp","landed_wt","tdate","grid"))
  melt_trips <- melt(port_trips, id.vars = c("veid","ftid","spid","tdate","grid"), measure.vars = "landed_wt")
  cast_trips <- dcast(melt_trips, ftid ~ spid, fun.aggregate = sum)
  rownames(cast_trips) <- cast_trips$ftid
  cast_trips$ftid <- NULL
  bc <- vegdist(cast_trips, method = "bray")
  bc_sim <- abs(bc-1)
  bc_mat <- as.matrix(bc_sim)
  
  bc_net <- graph.adjacency(bc_mat, mode="undirected", weighted = TRUE)
  ic <- infomap.community(bc_net, e.weights = E(bc_net)$weight)

  cg <- contract.vertices(bc_net, membership(ic))
  #E(cg)$weight <- 1 # leave this off, let the weights add up as they are. 
  cg2 <- simplify(cg, remove.loops=TRUE, edge.attr.comb = median) # take median similarity
```

```{r OXN_target_plot, echo=FALSE, warning=FALSE, message=FALSE}
V(cg2)$name <- c("red_urchin","rock_crab","lobs","cukes","chili","seabass","spot_prawn","rockfish","sheephead", "sable_ssp","oshrimp","angel_shark","copper_rockf","ridge_prawn", "scorp.fish","cabezon","sword","yellowtail","squid","albacore", "thds")

V(cg2)$size <- table(membership(ic))
l <- layout.fruchterman.reingold(cg2, niter=1000, area=vcount(cg2)^3, repulserad=vcount(cg2)^3)
par(mai = rep(0,4), bg="white")
plot(cg2, vertex.size = log(V(cg2)$size)*2.5, layout = l, edge.width = E(cg2)$weight*20, edge.label = "", vertex.label.font=1, vertex.label.color="black", vertex.frame.color="steelblue", vertex.color = "steelblue", edge.color=alpha("grey",1-E(cg2)$weight), vertex.label.family = "sans", vertex.label.cex = .75)

```

```{r bibartite_graph_OXN, echo=FALSE, message=FALSE, warning=FALSE}
# assign target cluster to each trip
trips <- port_trips
trips$target <- rep(NA, nrow(trips))
for(i in 1:length(communities(ic))){
  trips$target[which(trips$ftid %in% rownames(bc_mat[communities(ic)[[i]],]))] <- V(cg2)$name[[i]]
}

# now make bipartite network for gear
trip_types <- table(trips$target, trips$grid)
trip_gear <- melt(trip_types)
trip_gear <- subset(trip_gear, value > 10)
colnames(trip_gear) <- c("from","to","weight")

fisheries <- graph.data.frame(trip_gear, directed = FALSE)
E(fisheries)$weight <- trip_gear$weight
num_targets <- length(unique(trip_gear$from))
num_gears <- length(unique(trip_gear$to))
V(fisheries)$type <- c(rep(TRUE,num_targets), rep(FALSE, num_gears))
V(fisheries)$color <- c(rep("steelblue", num_targets), rep("brown", num_gears))
par(mai=rep(0,4), bg="white")
plot(fisheries, edge.width = E(fisheries)$weight/100, layout=layout.bipartite, ylim=c(-1,1),vertex.label.cex=.5, vertex.label.color="white", vertex.label.family = "sans", vertex.frame.color=V(fisheries)$color)
```

```{r participation_graph_OXN, echo=FALSE, message=FALSE, warning=FALSE}
# now make participation network. 
trips$fishery <- paste(trips$target, trips$grid, sep="_")
# restrict to fisheries used
used_fisherys <- unique(paste(trip_gear$from, trip_gear$to, sep="_"))
used_trips <- subset(trips, fishery %in% used_fisherys)
by_fishery <- table(used_trips$veid, used_trips$fishery)

fishery_mat <- matrix(ncol=length(used_fisherys), nrow=length(used_fisherys))
colnames(fishery_mat) <- colnames(by_fishery)
rownames(fishery_mat) <- colnames(by_fishery)

for(i in 1:ncol(by_fishery)){ # for each fishery except the last one
  fish_i <- colnames(by_fishery)[i]
  #cat("j: \n")
  for(j in i:ncol(by_fishery)){
    fish_j <- colnames(by_fishery)[j]
    num_ves = length(which(which(by_fishery[,i]>0) %in% which(by_fishery[,j] > 0)))
    
    mat_row <- which(rownames(fishery_mat)==fish_j)
    mat_col <- which(colnames(fishery_mat)==fish_i)
    
    fishery_mat[mat_row, mat_col] <- num_ves
    #cat(j," ")
  }
  #cat("\n")
}

g<-graph.adjacency(fishery_mat,weighted=T, diag=F, mode = "lower")
l <- layout.fruchterman.reingold(g, niter=500, area=vcount(g)^2, repulserad=vcount(g)^2)
#plot(g, edge.width = E(g)$weight, layout = l)

# looking for communities of vessels 

ic_gear <- infomap.community(g)
library(RColorBrewer)
paint <- brewer.pal(6,"Dark2")
V(g)$membership <- ic_gear$membership
V(g)[membership==1]$color <- paint[1]
V(g)[membership==1]$frame.color <- paint[1]
V(g)[membership==2]$color <-paint[2]
V(g)[membership==2]$frame.color <-paint[2]
V(g)[membership==3]$color <- paint[3]
V(g)[membership==3]$frame.color <-paint[3]
V(g)[membership==4]$color <-paint[4]
V(g)[membership==4]$frame.color <-paint[4]
V(g)[membership==5]$color <- paint[5]
V(g)[membership==5]$frame.color <- paint[5]
V(g)[membership==6]$color <- paint[6]
V(g)[membership==6]$frame.color <- paint[6]
par(mai=rep(0,4))

V(g)$size <- colSums(by_fishery)

plot(g, vertex.color=V(g)$color, edge.width = E(g)$weight, vertex.size=log(V(g)$size)*3, vertex.frame.color=V(g)$frame.color, vertex.label.color="black",layout=l, vertex.label.family = "sans", vertex.label.cex = .75)
```
