---
title: "Cluster_cp"
author: "Emma Fuller"
date: "April 29, 2014"
output: html_document
---

# Catch Profiles - all data

```{r load_data_packages, echo=FALSE,message=FALSE}

require(cluster)
require(plyr)
require(ggplot2)
require(data.table)
require(reshape2)
require(dplyr)
require(RColorBrewer)

load("/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/input_data/lb_tripTable_2014-05-01.Rdata")
load("/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/input_data/prop_tripTable_2014-05-01.Rdata")
load("/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/input_data/price_tripTable_2014-05-01.Rdata")

# set code parameters
max.clusts <- 30
samples = 100
sampsize = 1000
```

## PCA and Clustering log(lbs)
```{r loading,echo=FALSE}
# if data already exists, load that
file_load <- "/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_log.Rdata"
if(file.exists(file_load)){
  load(file_load)
} else{

# PCA
pca_loglb <- log_tripTable[[1]][,!"tripID", with=FALSE]
pca_out_loglb <- prcomp(pca_loglb, scale=TRUE)
npc <- length(which(summary(pca_out_loglb)[[6]][3,]<= .81))

dat_log <- pca_out_loglb$x[,1:npc]
  # randomize rows to prevent any bias from order of fish tickets
  data_log <- dat_log[sample(nrow(dat_log)),]
clust.dat.log <- vector("list",length = max.clusts)

# Clustering
set.seed(2)
for(i in 1:max.clusts){
  clust.dat.log[[i]] <- clara(dat_log, i, stand=TRUE, samples = samples, sampsize=sampsize, keep.data = FALSE, pamLike=TRUE, rngR = TRUE)
  print(i)
  }

objectives <- vector(length = max.clusts)
for (i in 1:max.clusts) { objectives[i] <- clust.dat.log[[i]]$objective }

asw <- vector(length = max.clusts)
for (i in 2:max.clusts) { asw[i] <- clust.dat.log[[i]]$silinfo$avg.width  }
max_asw <- max(asw)
nodes <- which(asw==max(asw))

# save relevant data
clust_log <- list(data_transform="log",cluster_sol = nodes, objectives = objectives, asw = asw, clustering = clust.dat.log[[nodes]]$clustering, sampsize=sampsize, samples=samples, npc=npc)

save(clust_log, file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_log.Rdata")
}

par(mfrow=c(1,2))
plot(clust_log$objectives, type="o",pch=19, cex=0.35,bty="n", main="Objective function")
plot(clust_log$asw, type="o", pch=19, cex=0.35, bty="n", main="Silhouette Width")
```

I retain `r clust_log$npc` principal components for clustering (which preserves > 80% of the cumulative variance). The maximum average silhouette width is `r max(clust_log$asw)`, and is found at the `r clust_log$cluster_sol` node solution. Examining catch composition for these two clusters I find that 

```{r plotting, cache=FALSE, echo=FALSE, warning=FALSE,fig.width=9, fig.height=13}
cp2 <- lb_tripTable[[1]]
cp2 <- cp2[,cp2 := clust_log$clustering]

# what are the mean catches of each species, in order of abundance?
melt_cp2 <- melt(cp2[,!"cp2",with=FALSE])
setkey(melt_cp2, "tripID")
melt_cp2_1 <- melt(cp2[,c("tripID", "cp2"), with=FALSE])
setnames(melt_cp2_1,names(melt_cp2_1),c("tripID","clustID","cluster2"))
melt_cp2_1 <- melt_cp2_1[,!"clustID",with=FALSE]
setkey(melt_cp2_1, "tripID")
melt_cp2 <- melt_cp2[melt_cp2_1]

  # median abundance by species
    median_abund <- melt_cp2[,as.double(median(value)),by="variable,cluster2"]
    sd_abund <- melt_cp2[,sd(value),by="variable,cluster2"]
    setnames(sd_abund,"V1","sd")
    median_abund <- median_abund[,"sd":= sd_abund[,"sd",with=FALSE]]

    ranked_abund <- ddply(median_abund, .(cluster2), transform, rank = rank(-V1, ties.method = "first"))
    top10 <- subset(ranked_abund, rank < 11)

    # now remove any top 10 that are =0, only there to fill out 10 slots
    top10 <- subset(top10, V1>0)

paint <- colorRampPalette(brewer.pal(8,"Dark2"))(length(unique(top10$variable)))
species.num <- as.numeric(as.factor(top10$variable))
ggplot(top10, aes(x=factor(variable), y=V1)) + 
  geom_bar(stat="identity",aes(fill=factor(variable))) + 
  facet_wrap(~cluster2,scale="free",ncol=3) + 
  theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, vjust=1)) 
  # mean abundance
    mean_abund <- melt_cp2[,mean(value),by="variable,cluster2"]
    ranked_abund <- ddply(mean_abund, .(cluster2), transform, rank = rank(-V1, ties.method = "first"))
    top10 <- subset(ranked_abund, rank < 11)

  # now remove any top 10 that are =0, only there to fill out 10 slots
    top10 <- subset(top10, V1>0)
ggplot(top10, aes(x=factor(variable), y=V1)) + 
  geom_bar(stat="identity",aes(fill=factor(variable))) + 
  facet_wrap(~cluster2,scale="free",ncol=3) + 
  theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, vjust=1))
```

## PCA and Clustering lbs
```{r, echo=FALSE,results='hide',eval=TRUE}

file_load <- "/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_lbs.Rdata"
if(file.exists(file_load)){
  load(file_load)
} else{

# PCA
pca_lb <- lb_tripTable[[1]][,!"tripID", with=FALSE]
pca_out_lb <- prcomp(pca_lb, scale=TRUE)
npc <- length(which(summary(pca_out_lb)[[6]][3,]<= .81))

dat_lb <- pca_out_lb$x[,1:npc]
  # randomize rows to prevent any bias from order of fish tickets
  data_lb <- dat_lb[sample(nrow(dat_lb)),]
clust.dat.lb <- vector("list",length = max.clusts)

set.seed(2)
for(i in 1:max.clusts){
  clust.dat.lb[[i]] <- clara(dat_lb, i, stand=TRUE, samples = samples, sampsize=sampsize, keep.data = FALSE, pamLike=TRUE, rngR = TRUE)
  print(i)
  }

objectives <- vector(length = max.clusts)
for (i in 1:max.clusts) { objectives[i] <- clust.dat.lb[[i]]$objective }

asw <- vector(length = max.clusts)
for (i in 2:max.clusts) { asw[i] <- clust.dat.lb[[i]]$silinfo$avg.width  }

max_asw <- max(asw)
nodes <- which(asw==max(asw))

# save relevant data
clust_lbs <- list(data_transform="lbs",cluster_sol = nodes, objectives = objectives, asw = asw, clustering = clust.dat.lb[[nodes]]$clustering, sampsize=sampsize, samples=samples,npc=npc)

save(clust_lbs, file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_lbs.Rdata")
    }
    
par(mfrow=c(1,2))
plot(clust_lbs$objectives, type="o",pch=19, cex=0.35,bty="n", main="Objective function")
plot(clust_lbs$asw, type="o", pch=19, cex=0.35, bty="n", main="Silhouette Width")

```

I retain `r clust_lbs$npc` principal components for clustering (which preserves > 80% of the cumulative variance). The maximum average silhouette width is `r max(clust_lbs$asw)`, and is found at the `r clust_lbs$cluster_sol` node solution. Examining catch composition for these clusters I find that 

```{r, echo=FALSE, warning=FALSE,eval=FALSE}
cp <- lb_tripTable[[1]]
cp <- cp[,"cp" := clust_lbs$clustering]

# what are the mean catches of each species, in order of abundance?
  melt_cp <- melt(cp[,!"cp",with=FALSE])
  setkey(melt_cp, "tripID")
  melt_cp_1 <- melt(cp[,c("tripID", "cp"), with=FALSE])
  setnames(melt_cp_1,names(melt_cp_1),c("tripID","clustID","cluster"))
  melt_cp_1 <- melt_cp_1[,!"clustID",with=FALSE]
  setkey(melt_cp_1, "tripID")
  melt_cp <- melt_cp[melt_cp_1]

  # median abundance by species
    median_abund <- melt_cp[,as.double(median(value)),by="variable,cluster"]
    sd_abund <-     melt_cp[,sd(value),by="variable,cluster"]
    setnames(sd_abund,"V1","sd")
    median_abund <- median_abund[,"sd":= sd_abund[,"sd",with=FALSE]]

    ranked_abund <- ddply(median_abund, .(cluster), transform, rank = rank(-V1, ties.method = "first"))
    top10_med <- subset(ranked_abund, rank < 11)
    colnames(top10_med)[c(3,5)] <- c("median","rank_median")

    # now remove any top 10 that are =0, only there to fill out 10 slots
    #top10_med <- subset(top10_med, V1>0)

  # mean abundance
    mean_abund <- melt_cp[,mean(value),by="variable,cluster"]
    ranked_abund <- ddply(mean_abund, .(cluster), transform, rank = rank(-V1, ties.method = "first"))
    top10_mean <- subset(ranked_abund, rank < 11)
    colnames(top10_mean)[c(3,4)] <- c("mean","rank_mean")

  # now remove any top 10 that are =0, only there to fill out 10 slots
    #top10_mean <- subset(top10_mean, V1>0)

par(mfrow=c(1,2))
barplot(top10_med$median, col=top10_med$cluster,names.arg =top10_med$variable, las=2,main="median abundance")
barplot(top10_mean$mean, col=top10_mean$cluster,names.arg =top10_mean$variable, las=2, main="mean abundance")

# looking at it together
  top10_all <- merge(top10_med, select(top10_mean,variable,mean,rank_mean,cluster),by=c("variable","cluster"),all.x=TRUE,all.y=TRUE) 

```

Looking at median catch, there's many fewer species that zero medians, but non-zero means. This just means that the mean catches are extremely skewed, and that there are very few of them. This result is slightly better than the log(lb) run becaues it reduces that peak at clusters=2, but the 6 cluster solution doesn't immediately look right. Trying with price data

## PCA and Clustering price
```{r price_cluster, echo=FALSE,results='hide',eval=TRUE}

file_load <- "/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_price.Rdata"
if(file.exists(file_load)){
  load(file_load)
} else{

# PCA
pca_price <- price_tripTable[[1]][,!"tripID", with=FALSE]
pca_out_price <- prcomp(pca_price, scale=TRUE)
npc <- length(which(summary(pca_out_price)[[6]][3,]<= .81))

dat_price <- pca_out_price$x[,1:npc]
  # randomize rows to prevent any bias from order of fish tickets
  data_price <- dat_price[sample(nrow(dat_price)),]
clust.dat.price <- vector("list",length = max.clusts)
# Clustering
set.seed(2)
for(i in 1:max.clusts){
  clust.dat.price[[i]] <- clara(dat_price, i, stand=TRUE, samples = samples, sampsize=sampsize, keep.data = FALSE, pamLike=TRUE, rngR = TRUE)
  print(i)
  }

objectives <- vector(length = max.clusts)
for (i in 1:max.clusts) { objectives[i] <- clust.dat.price[[i]]$objective }

asw <- vector(length = max.clusts)
for (i in 2:max.clusts) { asw[i] <- clust.dat.price[[i]]$silinfo$avg.width  }

max_asw <- max(asw)
nodes <- which(asw==max(asw))

# save relevant data
clust_price <- list(data_transform="price",cluster_sol = nodes, objectives = objectives, asw = asw, clustering = clust.dat.price[[nodes]]$clustering, sampsize=sampsize, samples=samples)

save(clust_price, file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_price.Rdata")
}

par(mfrow=c(1,2))

plot(clust_price$objectives, type="o",pch=19, cex=0.35,bty="n", main="Objective function")
plot(clust_price$asw, type="o", pch=19, cex=0.35, bty="n", main="Silhouette Width")

```

I retain `r clust_price$npc` principal components for clustering (which preserves > 80% of the cumulative variance). The maximum average silhouette width is `r max(clust_price$asw)`, and is found at the `r clust_price$cluster_sol` node solution. Examining catch composition for these clusters I find that 

```{r price_plot, cache=FALSE, echo=FALSE, warning=FALSE,eval=FALSE}
cp <- price_tripTable[[1]]
cp <- cp[,"cp" := clust_price$clustering]

# what are the mean catches of each species, in order of abundance?
melt_cp <- melt(cp[,!"cp",with=FALSE])
setkey(melt_cp, "tripID")
melt_cp_1 <- melt(cp[,c("tripID", "cp"), with=FALSE])
setnames(melt_cp_1,names(melt_cp_1),c("tripID","clustID","cluster"))
melt_cp_1 <- melt_cp_1[,!"clustID",with=FALSE]
setkey(melt_cp_1, "tripID")
melt_cp <- melt_cp[melt_cp_1]

  # median abundance by species
    median_abund <- melt_cp[,as.double(median(value)),by="variable,cluster"]
    sd_abund <-     melt_cp[,sd(value),by="variable,cluster"]
    setnames(sd_abund,"V1","sd")
    median_abund <- median_abund[,"sd":= sd_abund[,"sd",with=FALSE]]

    ranked_abund <- ddply(median_abund, .(cluster), transform, rank = rank(-V1, ties.method = "first"))
    top10_med <- subset(ranked_abund, rank < 11)
    colnames(top10_med)[c(3,5)] <- c("median","rank_median")

    # now remove any top 10 that are =0, only there to fill out 10 slots
    #top10_med <- subset(top10_med, V1>0)

par(mfrow=c(1,2))

barplot(top10_med$median, col=top10_med$cluster,names.arg =top10_med$variable, las=2,main="median abundance")

  # mean abundance
    mean_abund <- melt_cp[,mean(value),by="variable,cluster"]
    ranked_abund <- ddply(mean_abund, .(cluster), transform, rank = rank(-V1, ties.method = "first"))
    top10_mean <- subset(ranked_abund, rank < 11)
    colnames(top10_mean)[c(3,4)] <- c("mean","rank_mean")

  # now remove any top 10 that are =0, only there to fill out 10 slots
    #top10_mean <- subset(top10_mean, V1>0)

barplot(top10_mean$mean, col=top10_mean$cluster,names.arg =top10_mean$variable, las=2, main="mean abundance")

# looking at it together
top10_all <- merge(top10_med, select(top10_mean,variable,mean,rank_mean,cluster),by=c("variable","cluster"),all.x=TRUE,all.y=TRUE) 

```

Looking at median prices, we're back to two species and mud shrimp are consistently bringing in dollars.  This again just means that the mean catches are extremely skewed, and that there are very few of them. This result is just like the results for log(lbs). 

## PCA and Clustering proportion

```{r cluster_prop, echo=FALSE,results='hide',eval=TRUE}
file_load <- "/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_prop.Rdata"
if(file.exists(file_load)){
  load(file_load)
} else{
pca_prop <- prop_tripTable[[1]][,!"tripID", with=FALSE]
pca_out_prop <- prcomp(pca_prop, scale=TRUE)
npc <- length(which(summary(pca_out_prop)[[6]][3,]<= .81))

dat_prop <- pca_out_prop$x[,1:npc]
  # randomize rows to prevent any bias from order of fish tickets
  data_prop <- dat_prop[sample(nrow(dat_prop)),]
clust.dat.prop <- vector("list",length = max.clusts)

set.seed(2)
for(i in 1:max.clusts){
  clust.dat.prop[[i]] <- clara(dat_prop, i, stand=TRUE, samples = samples, sampsize=sampsize, keep.data = FALSE, pamLike=TRUE, rngR = TRUE)
  print(i)
  }

objectives <- vector(length = max.clusts)
for (i in 1:max.clusts) { objectives[i] <- clust.dat.prop[[i]]$objective }

asw <- vector(length = max.clusts)
for (i in 2:max.clusts) { asw[i] <- clust.dat.prop[[i]]$silinfo$avg.width  }

max_asw <- max(asw)
nodes <- which(asw==max(asw))

# save relevant data
clust_prop <- list(data_transform="prop",cluster_sol = nodes, objectives = objectives, asw = asw, clustering = clust.dat.prop[[nodes]]$clustering, sampsize=sampsize, samples=samples)

save(clust_prop, file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_prop.Rdata")

}

par(mfrow=c(1,2))
plot(clust_prop$objectives, type="o",pch=19, cex=0.35,bty="n", main="Objective function")
plot(clust_prop$asw, type="o", pch=19, cex=0.35, bty="n", main="Silhouette Width")
```

I retain `r clust_prop$npc` principal components for clustering (which preserves > 80% of the cumulative variance). The maximum average silhouette width is `r max(clust_prop$asw)`, and is found at the `r clust_prop$cluster_sol` node solution. Examining catch composition for these clusters I find that 

```{r, echo=FALSE, warning=FALSE,eval=TRUE,fig.width=3,fig.height=10}
cp <- prop_tripTable[[1]]
cp <- cp[,"cp" := clust_prop$clustering]

# what are the mean catches of each species, in order of abundance?
melt_cp <- melt(cp[,!"cp",with=FALSE])
setkey(melt_cp, "tripID")
melt_cp_1 <- melt(cp[,c("tripID", "cp"), with=FALSE])
setnames(melt_cp_1,names(melt_cp_1),c("tripID","clustID","cluster"))
melt_cp_1 <- melt_cp_1[,!"clustID",with=FALSE]
setkey(melt_cp_1, "tripID")
melt_cp <- melt_cp[melt_cp_1]

  # median abundance by species
    median_abund <- melt_cp[,as.double(median(value)),by="variable,cluster"]
    sd_abund <-     melt_cp[,sd(value),by="variable,cluster"]
    setnames(sd_abund,"V1","sd")
    median_abund <- median_abund[,"sd":= sd_abund[,"sd",with=FALSE]]

    ranked_abund <- ddply(median_abund, .(cluster), transform, rank = rank(-V1, ties.method = "first"))
    top10_med <- subset(ranked_abund, rank < 11)
    colnames(top10_med)[c(3,5)] <- c("median","rank_median")

ggplot(top10_med, aes(x=factor(variable), y=median)) + geom_bar(stat="identity",aes(fill=factor(variable))) + facet_wrap(~cluster,scale="free_x",ncol=3)

    # now remove any top 10 that are =0, only there to fill out 10 slots
    #top10_med <- subset(top10_med, V1>0)

  # mean abundance
    mean_abund <- melt_cp[,mean(value),by="variable,cluster"]
    ranked_abund <- ddply(mean_abund, .(cluster), transform, rank = rank(-V1, ties.method = "first"))
    top10_mean <- subset(ranked_abund, rank < 11)
    colnames(top10_mean)[c(3,4)] <- c("mean","rank_mean")

  # now remove any top 10 that are =0, only there to fill out 10 slots
    #top10_mean <- subset(top10_mean, V1>0)

ggplot(top10_mean, aes(x=factor(variable), y=mean)) + geom_bar(stat="identity",aes(fill=factor(variable))) + facet_wrap(~cluster,scale="free_x",ncol=3)

# looking at it together
top10_all <- merge(top10_med, select(top10_mean,variable,mean,rank_mean,cluster),by=c("variable","cluster"),all.x=TRUE,all.y=TRUE) 
```
