---
title: "draft"
author: "Emma Fuller"
date: "March 5, 2015"
output: pdf_document
---

# Question

I seek to describe how individual income variability is related to local ecological, economic and management related characteristics. The goal of this analysis is to discover if market accessibility, habitat availability, and/or access to open-access fisheries is related to income variability. 

# Motivation (conservation)

+ Stakeholder wellbeing is an increasingly important measure of conservation success (i.e. ecosystem services, etc.) and is a formal one for US fisheries ("fishing community health")
+ We examine income variability as a measure of human wellbeing. 
+ Natural resource-based livlihoods (fisheries, farming, etc.
) can experience income variability for any of three reasons: environmental (vagaries of weather, disease, abundance of resource), economic (market access, demand, price), or management (restricted access).
+ Previous work shows that diversity of fisheries participation is negatively related to income variability and suggests that geography and management may play large roles in determining how diverse income is. 
+ We add to this literature by replicating diversity-income variability relationship, unpack the proxy variable of geography and explicitly test to what extent ecology, markets and management can explain this trend in income variability. 

# Methods
Data manipulations

+ Income
    + Subset to vessels we surmise make full income off of fishing (median income > $10,000)
    + Calculate income volatility (interannual variation - coefficient of variation; min/mean - measure of extreme risk, **not-done**)
+ Fisheries participation
    + Determine fishery to which each trip belongs (metiers)
    + Calculate diversity of fisheries participation (simpson index of revenue)
    + Find fishing strategies: cluster based on proportion of annual revenue in each fishery by vessel
+ Location
    + Calculate median latitude of catch (**not done**)

# Results
After clustering fisheries participation (by proportion of revenue from each fishery), I get 29 different "fishing strategies" or typical combinations of fisheries in a year. Each vessel can be a member of a new fishing strategy each year, but about 3/4 of the vessels stay in the same fishing strategy across all five years. 

Below I plot the distributions of income variablity (coefficient of variation of annual revenue) by fishing strategy. If a vessel did not participate in the same strategy across all 5 years, they are thrown in the "multi" category.[^6] I also show where the variation in revenue is equivalent to the average expected revenue ($cov = 1$). Based on this, it looks like some strategies do better than others. 

[^6]: the reason there are some blank rows are fishing strategies that exist, but no vessel does that one strategy across all five years.  

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.width=9}
library(pander)
yrdf <- readRDS("/Users/efuller/1/CNH/Analysis/Metiers/writing/code/1_cleaningData/yrdf.RDS")

bymedian <- with(yrdf[[2]], reorder(x=single_cluster, -cv_adj_revenue, median))
boxplot(cv_adj_revenue~ bymedian, yrdf[[2]], col = 'grey', ylab="coefficient of variation (annual income)", xlab="fishing strategy")
abline(h=1, lwd=3,lty=2, col="indianred")
```

Fitting a multiple regression, I find that if I add fishing strategy ID (cluster), that some fishing strategies are associated with a decrease in mean variability, while a few an increase, and many that have no effect at all. I drop the quadratic diversity term in this model, but I find that the main diversity term is negative and still significant. Which means that regardless of what fishing strategy you choose, the more diverse you are, the lower we expect the annual variation.

```{r, echo=FALSE,warning=FALSE,message=FALSE}
lm_pp <- lm(cv_adj_revenue ~ mean_simpson + single_cluster, data = yrdf[[2]])
summary(lm_pp)
```

Interestingly the intercept (I think) is equivalent to category 1, which is the fishing strategy that features more than 90% of your participation in dungeness crab. This has a hugely positive effect (`r round(lm_pp$coefficients[1],3)`), which means that your cv(revenue) goes up by around .6 if you participate in crab, regardless of how diverse you are. 

I also find that this model is a better fit (according to AIC)

```{r,warning=FALSE,message=FALSE}
lm_reg <- lm(cv_adj_revenue ~ mean_simpson + I(mean_simpson^2), data = yrdf[[2]])
AIC(lm_reg) - AIC(lm_pp)
```

And a better fit from the most basic model
```{r}
lm_simp <- lm(cv_adj_revenue ~ mean_simpson, data = yrdf[[2]])
AIC(lm_simp) - AIC(lm_pp)
```

# Discussion
This suggests to me that the fisheries in which you participate are important, but that also diversity itself is still helpful. I'm not quite sure how to structure the next step. Some possible questions:

+ Why doesn't everyone do the successful fishing strategies? Does management block people from participating in those fishing strategies? Are those fisheries that make up "good" fishing strategies limited entry?
+ Diversity itself is still important, regardless of what fishing strategy you undertake. And because some fisheries are going to have to stay limited entry, knowing whether we could improve diversity of participation by other means[^7] would be vaulable. So what explains diversity?

[^7]: i.e. through restoration of species/habitat or market structure

# Appendix: choosing indicators?


For each vessel it will be the weighted catch average of ports to which vessels deliver
 
## Ecology predictors 
 
+ proportion area of fishable (i.e. non protected) area in $X$ radius of port. (is there a large MPA blocking most near-shore fishermen?)[^1]
+ proportion area of rock in $X$ radius of port (heard that "rocks" were crucial to being able to do near-shore rockfish assemblages)[^2]
+ proportion of rock habitat protected in $X$ radius of port (no good if that habitat is all protected. Alternatively it could be positively associated due to spillover)[^3]
+ proportion of rock habitat unprotected in $X$ radius of port (reciprocal of above)
+ do similar with "hard" habitat? or other types of habitat (mixed, soft)

[^1]: overlay MPA polygons over radius of port, calculate proportion of area fishable
[^2]: get habitat layers from essential habitat analysis, calculate proportion of area with rock habitat
[^3]: combine MPA polygons and habitat layers, calculate proportion of areas protected and rocky


## Management predictors

+ TBD metrics, maybe the proportion of fisheries that are open access

## Market predictors

+ distance to major city (how to define -- i.e. Seattle, Portland, SF, LA)
+ number of first recievers (recievers are proxies for processors, more recievers, more options?)
+ evenness of first recievers revenues (one processor effectively dominating?)