---
title: "People in EBM"
author: "Emma Fuller"
date: "January 3, 2015"
output:
  pdf_document:
    keep_tex: yes
bibliography: refs.bib
---

# Introduction
We perform a metier analysis on the US west coast commercial fisheries. We demonstrate how this analysis improves upon previously used definitions of fisheries in the US, and show the utility of these definitions by building two different types of networks: a typical foodweb, bipartite network between fisheries and species fished; and a social-ecological network in which human linkages across fisheries are included. We then apply network measures to demonstrate how management-revelent social and ecological conclusions can be drawn from this analysis. In particular, we highlight how this approach can be useful to better include humans in ecosystem-based management. 

# Methods
We do a metier analysis using infoMap [more detail to come] and examine the resulting fisheries-deliniations.

We build a bipartite network using the metiers where the interaction strength between metier and species is the volume of species caught averaged over the 5 year dataset. 

We build a social-ecological network by adding a participation network to the bipartite foodweb. The participation network has as its nodes, metiers. The strength of connection between the fisheries is the number of individuals which participated in both fisheries averaged over 5 years. 

We first analyze the bipartite network to try to determine the following types of inferences

- vulnerability: highlight which species are more or less vulnerable (i.e. targeted by more than one fishery)
- generality: highlight which fisheries catch more or less species 

And then analyze the social-ecological network and show how many of the same metrics can be applied to an SES network.

- centrality: which fisheries are important for everyone (i.e. refuge fisheries)

# Results
```{r,echo=FALSE,warning=FALSE,message=FALSE}
tickets <- readRDS("code/3_exploreBuildwebs/tickets.RDS")
trips <- readRDS("code/3_exploreBuildwebs/trips.RDS")
library(e1071)

```

## This analysis improves upon heuristic definitions of fisheries
Commonly suggested heuristics to define species targeted often center on species that are the majority of a trip's catch either by weight or revenue. We compare our definitions of fisheries and find it reduces the number of fisheries by about 80% (from ~500 to ~100). 

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.width=7,fig.height=7}
library(vegan); library(RColorBrewer)
# rows are fisheries, columns are by revenue
percents <- table(trips$metier, trips$m.rev)/rowSums(table(trips$metier,trips$m.rev))
dca <- decorana(percents)
colnames(percents) <- tolower(colnames(percents))
rownames(percents) <- tolower(rownames(percents))
tabasco(decostand(percents,"log"), dca,col=c("white","white",rev(brewer.pal(9,"Blues"))),cexRow=.15,cexCol=.5, add.expr=list(abline(v=7,col="grey"), text(x = 15, y = 375, "tls_1\nsalmon troll")),xlab="metier-based definition", ylab="revenue-based definition")
```

Here I plot a matrix where columns are the metiers, and rows are fisheries defined as major species by revenue. The cell shows the $\log$ number of trips in which that species was the majority catch by revenue.[^1] Darker colors indicate many trips with this species as majority, lighter colors indicate relatively few species as a majority. The metiers are ordered using a detrended correspondance analysis (`decorana` function in the `vegan` package to aid in visualization, `R`). The plot demonstrates the reduction in dimensionality as a number of rare species are grouped into a single fishery. 

[^1]: Using the Adjusted Rand Index ($ARI$) I find no difference between classification using volume or revenue ($ARI=$ `r classAgreement(table(trips$m.vol, trips$m.rev))$crand`)

## This analysis allows building of food webs that include humans
Ecosystem based management strives to manage an ecosystem holistically, and include humans in the analysis [@larkin1996concepts; @lucey2013operational]. Here we demonstrate such a foodweb. 

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.width=8}
library(bipartite)
web <- readRDS("code/3_exploreBuildwebs/web.RDS")
# take top 10 by volume
colnames(web) <- tolower(colnames(web))
rownames(web) <- tolower(rownames(web))
# plotweb(web[,names(sort(colSums(web),decreasing=T)[1:10])],arrow="down",labsize = .5)

# try removing any that have fewer than 100 pounds across all five years
web[which(web<=100)]=0
plotweb(web[,names(sort(colSums(web),decreasing=T)[1:10])],arrow="up",labsize = .5)
```
Here I plot a bipartite network in which upper nodes are metier-defined fisheries and the lower nodes are fish species. The strength of interaction is to the total volume harvested over the 5 year dataset. For visual clarity, this network is retricted to showing the top 10 fisheries by volume and the species harvested. Further the species caught are restricted to those which had at least an average of 20 lbs caught per year. 

Using this network, we can estimate the vulnerability of species by the number of seperate fisheries which exploit them, and the intensity with which they do so. 

```{r, echo=FALSE,warning=FALSE,message=FALSE}
# see networklevel() in bipartite for other analyses - specifically generality and vulnerability
networklevel(web[,names(sort(colSums(web),decreasing=T)[1:10])], index="vulnerability")
```

## Can improve this web by considering human-mediated connectivity between fisheries

# References
