---
title: 'You can only manage what you measure: A data-driven evaluation of ocean policies
  for vibrant coastal communities'
author: "Emma Fuller"
date: "July 18, 2015"
output:
  word_document: default
  pdf_document:
    number_sections: yes
  html_document: default
csl: the-american-naturalist.csl
bibliography: method_refs.bib
---

# Question
How to evaluate the impact of a change in ecosystem properties (abiotic, biotic, and/or management driven) on the coastal communities dependent on the ecosystem? Here we use the implementation of ITQS to demonstrate a data-driven method to evaluate of the impact of management change on coastal port fishing communities. 

# Motivation
[p1]

+ Increasingly natural resource management is mandated to consider human wellbeing alongside the integrity of ecosystem processes and functions.
+ Management is also tasked with setting explicit and measurable goals 
+ We lack ways measuring human wellbeing, especially at the community level. 

[p2]

+ This gap in methods for quantifying change in human communities makes it difficult to set explicit mangement goals. 
+ This is especially true of fisheries, which are mandated by the National Ocean Policy to not only
_[d]etermine the impacts of interacting stressors on ecological systems, economies, and communities_ but to also develop _[e]ffective management of activities that affect ocean health_ which will _require considering ... how human activities impact both the ecosystems themselves, and the communities that depend on them._

[p3]

+ Previous fisheries work which considers the impacts to fishing communities most often calculates individual economic/demographic impacts and aggregates up
+ Other work have expanded the scope to consider how impacts affect entire supply chains from fishermen, processors, to others tangentially involved in the commercial fsihery. 
+ While useful, the drawback of these approaches is that it simplifies fishers to effectivley participating in only one fishery.  
+ Instead, we know that in many places fishermen participate in multiple fisheries, and so impacts in one fishery may ripple more widely to change that harveter's participation in other fisheries. 

[p4]

+ Here we provide a method to quantify the impact of a change in management across port communities along the US west coast. 
+ To do so, we first provide a data-driven method to define fisheries concurantly from one another, a gap that exists given the traditional focus of single fisheries management. 
+ This approach unites the individaul-level focus common in fisheries economics to the ecosystem-based management approach popular among marine ecologists, conservationists and fisheries anthropologists. Because community-level language is common in federal policy, we belive this will be a particularly useful method to use in evaluating community level responses to management change.

# Materials and Methods
This work makes use of landings data and vessel characteristics to classify landings into fisheries. We use these fisheries designations combined with port landings to characterize port communities' fisheries participation. Finally we examine how port communities change before after ITQ implementation in the groundfish trawl fishery. All analyses are performed using `R` [@R2015] unless otherwise stated.

## Data preparation
Fisheries landings data comes from the Pacific Fishery Information Network database [@pacfin2014], and include commercial landings from all vessels fishing in the US Extended Economic Zone (EEZ) off the US west coast between 2009 and 2013. Landings receipts report the amount of fish caught by market category. These market categories are approximately equivalent to species-level identification, although not exact. Rockfish are most likely to be approximate, as species are often difficult to distinguish between. However, the bigger the price differential between species, the more accurate the market categories.[^1] Thus in the following work we assume that vessels are targeting market categories, rather than species, as that is the economically relevant determination. 

[^1]: Personal communication from Brad Stenberg

For each market category, landing tickets report price per pound, date, port of landing, and vessel identifying information. After adjusting for inflation using the 2009 Consumer Price Index, we calculate vessel revenues by multiplying the price per pound by the number of pounds landed.[^2] We include only vessels with average annual revenues above $10,000 (adjusted to 2009 values), to exclude vessels for which fishing was a part time activity. To focus on major species, we remove landings from species that show up in fewer than 100 trips across our five year dataset and whose median catch is less than 100 pounds.[^3] 

[^2]: This is currently `ppp*landedwt`, but maybe should be using `roundwt`. Need to check

[^3]: Am going to try the clustering without this, I think this was based on trying to do k-means and is no longer necessary. Also it's an awkward thing to have to defend. And then when we're looking at species diversity in landings data, we want to include those species again. So it seems overly unclear. 

## Defining fisheries
We define a fishery as a gear-type targetting a coherent species assemblage [@vanPutten:2011bj]. The Pacific Management Council has developed a set of sector based definitions similar to this approach for the federally managed groundfish landings, but no equivalent exists for non-groundfish fisheries. In order to treat the landings dataset uniformly, we apply a metier-like analysis to this landing data [@Deporte:2012kq]. Our goal was to develop a method that could be used across fisheries with wildly different assemblages (i.e. distinguishing rockfish assemblages from groundfish trawls from pink shrimp trawls) in an unsupervised manor. Previous metier analyses occur largely in Europe, although metier-like analyses have been performed in the Northeast US, classifying fishing data to define “operational fisheries” of New England [@Lucey:2013jy]. A metier is defined as a gear-species target(s) combination [@Deporte:2012kq]. The first step of these methods is to define species targets through clustering species composition of landings. 

There exists number methods by which to find characteristic assemblages in data, and is a topic that's recieving much attention in the 'big data' revolution. Previous metier analyses have used a variety of clustering algorithms including k-means [@Lucey:2013jy], hierarchical algomorative clustering [@Deporte:2012kq], and hierarchical ascending classification [@Pelletier:ISmLZefX]. Catch data is a high dimensional dataset, due to the fact that a single groundfish trawl can bring in > 20 species. Because these clustering methods do poorly with high-dimensional data [CITE], dimensionality reducing approaches in the form of PCAs are often applied prior to the clustering step.

Because our goal is to develop a single analysis pipeline to classify fisheries coast-wide (rather than between a few mixed fisheries as has been done previously), our dimensionality was three times the dimensionality of previously analyzed landings data.  With this degree of complexity in our data, ordination techniques were not sufficent to reduce the data to two to three principal axes for which k-means and hierarchical clustering do best. Finally, because we desired an unsupervised approach to classify fisheries, we avoided exploratory methods such as k-means and hierarchical clustering which require a subjective descision of an optimal clustering solution [CITE - also BIC tecniques for cluster solutions]. 

Given our goals, we use the infoMap algorithm. infoMap is an information theoretic approach, which uses the probability flow of a random walker on a network as a proxy for the information flows in a real system. The objective of infoMap is to compress the description of the probability flow, and in doing so partitions the network into modules. infoMap works by computing the fraction of time a node is visited by a random walker using a deterministically greedy search algorithm. Merges between modules that give the largest decrease in description length are made until further merging leads to increases of description length. Results are refined with a simulated annealing approach, starting at several different temperatures, with the run selected as the one that gives the shortest description of the network.

### Clustering pipeline
To find species target data we classify target species assemblages by first subsetting to all 2010 trips and searching for characteristic catch compositions. To find these assemblages we first split trips by gear type (using PacFIN `grgroups` designation, see table below).

```{r gear.groups.table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
setwd("/Users/efuller/1/CNH/")

# if file doesn't exist, download from PacFin
  if(!file.exists("rawData/Catch/gr_tree.txt")){
    system('wget -O rawData/Catch/gr_tree.txt "http://pacfin.psmfc.org/pacfin_pub/data_rpts_pub/code_lists/gr_tree.txt"')
  }

# load gear table
  grid <- read.fwf("/Users/efuller/1/CNH/rawData/Catch/gr_tree.txt", widths = c(22,54), skip = 7, n = 65, stringsAsFactors = FALSE)
  colnames(grid) <- c("gear.group","description")
  library(dplyr); library(stringr)
  # cleaning up the frame, remove empty lines, lower case, remove columns, trim whitespace
  grid <- grid[complete.cases(grid),]; 
  grid <- grid %>%
    mutate_each(funs(tolower)) %>%
    mutate_each(funs(str_trim))
  
  grid <- grid[grep("^all", grid$description),] # just get gear groups
  grid$gear.group <- gsub('\\\\_',"",grid$gear.group) # remove the weird slashes

# load cleaned FTL data and calculate number of vessels and trips for each gear group
  ftl <- readRDS("/Users/efuller/1/CNH/processedData/catch/1_cleaningData/filtered_ftl.RDS")
  
  # tabulate
  gear.tab <- ftl %>%
    group_by(grgroup) %>%
    summarize(n.vessels = length(unique(drvid)), 
              n.trips = length(unique(trip_id))
              ) %>%
    mutate(grgroup = tolower(grgroup)) %>%
    arrange(desc(n.vessels)) %>%
    rename(gear.group = grgroup)
  
  # put together
  gear.tab <- merge(gear.tab, grid, by = "gear.group",all.x = TRUE, all.y = FALSE, sort = FALSE)
  
  # display table
  library(pander)
  set.alignment('left', row.names = 'left')
  set.caption("PacFin gear groups, participation, and descriptions for US west coast landings data 2009-2013")
  panderOptions('table.split.table', Inf)
  pander(gear.tab, style = 'rmarkdown')
```

We calculate a pairwise dis-similiarity index for each trip within a gear/year subset using the Bray-Curtis dissimilarity index. This metric has the advantage of avoiding the double-0 problem common in species count data. The Bray-Curtis dissimilarity index is defined as
$$BC_{ij}=\frac{2C_{ij}}{S_i + S_j}$$
where $C_ij$ is the biomass of the lesser value for only those two species in common between both sites. $S_i$ and $S_j$ are the total number of individuals counted in both trips. This index ranges between zero and one, with zero meaning the sites have the same composition and one meaning they share no species.

We transform the dissimilarity index to be a measure of similarity
$$\text{Similarity} = 1 - \mid BC \mid$$
and build an undirected, weighted network in which nodes are trips, and edge widths are the similarity in species composition between trips. This allows a vessel to be represented in multiple nodes if it makes trips that vary substantially in catch composition. With this network we use the infoMap algorithm to find communities (clusters or subgraphs) within the network (Martin Rosvall and Bergstrom 2008; M Rosvall, Axelsson, and Bergstrom 2009).

After dropping any modules that have fewer than five trips, we use a k-nearest-neighbor (knn) classifier to assign all other trips of each gear subset to those possible metiers. The nearest neighbor to each trip was found using the Bray-Curtis dissimilarity index (transformed into a similarity). 

## Measuring individual impacts
To measure the change in the number of fisheries a vessel participates in we calculate the Simpson's index for each vessel each year where number of trips in each metier is the unit of diversity. Using these diversity scores, we compare a vessel's participation diversity in the period before and after catch shares. Thus vessels had to land at least $10,000 worth of fish in either (or both of) 2009, 2010 and $10,000 in either (or both) or 2012, 2013. If both years were available in the before and after, we took the average of the annual diversity scores.

We measure the response variable as both the delta and the ratio between post and pre catch share diversity for each vessels. Treatment is whether or not the vessel ever landed a landing using in 2011 and/or 2012. 

## Measuring community impacts 
We introduce the concept of participation networks. Here the nodes are fisheries, and the strength of the edges between them are proportional to the number of vessels which participate in the same fishery in some time window. Specifically, edge weight is measured as the bray curtis similarity between the fleet composition in fishery $i$ and fishery $j$.

These networks allow the application of existing network measures to characterize how participation varies across communities of fishers. In this case, we consider the port the unit of community, based in part on the previous work NOAA has done in characterizing port communities. 

Here we use the implementation of ITQs for the west coast groundfish trawl fishery as a natural experiment to see how it affected communities. Previous work has hypothesized that ITQs will allow vessels to be more flexible, since they face less of an opportunity cost with ITQs. We measured both the individual level of diversity for vessels which fished in the groundfish ITQ fishery before and after ITQ implementation. We also examined the changes in network statistics. By correlating the invidual fleet-specific changes with those observed in hte larger network we demonstrat the far reaching changes caused by ITQs. 

# Results
## Defining fisheries
In our analysis temporal and spatial structure emerges from the data, and we are able to recover the commonly recognized major fisheries and their seasonality, along with more spatially and temporally restricted fisheries. Our methods have the additional benefit of only requiring the catch composition of trips, making it possible to integrate data from both state and federal management databases which lack consistent permitting data across states.

## Participation Networks
The first observation of the participation networks is how visually distinct each appears in structure and composition. 

We find that participation networks compliment individual-level characterizations of fishing behavior. Specifically the diversity of participation, a characteristic shown to reduce revenue variability at the vessel level, is determined by the number and evenness with which a vessel participates in available fisheries. Our participation networks highlight that the same level of diversity can be observed in two communities despite one participation network being much more "diverse" (i.e. SB versus Newport).

## Individual impacts of management change
```{r echo=FALSE, warning=FALSE, message=FALSE}
ftl <- readRDS("/Users/efuller/1/CNH/processedData/catch/5_add_length/ftl_len.RDS")
yrdf <- readRDS("/Users/efuller/1/CNH/processedData/catch/1_cleaningData/yrdf.RDS")

library(dplyr)

ifqY <- unique(ftl$drvid[which(ftl$ifq_landing=="Y")])

# average vessel diversity before and after - vessels must be present in all years 2009-2013 
div_baci <- yrdf[[1]] %>%
  group_by(drvid) %>%
  mutate(all.year = ifelse(all(c(2009,2010,2012,2013) %in% unique(year)), 1, 0),
         no.div = ifelse(sum(simpsons)==0, 1, 0)) %>%
  filter(year != 2011, all.year == 1, no.div == 0) %>%
  select(drvid, simpsons, year, all.year) %>%
  group_by(year) %>%
  mutate(before = ifelse(year %in% c(2009,2010), 1, 0)) %>%
  group_by(drvid, before) %>%
  summarise(mean.simp = mean(simpsons, na.rm = T)) %>%
  group_by(drvid) %>%
  summarise(delta.simpson = diff(mean.simp))

div_baci$ifq <- ifelse(div_baci$drvid %in% ifqY, 1, 0)
div_baci$multi.port <- ifelse(div_baci$drvid %in% yrdf[[2]]$drvid[which(yrdf[[2]]$single_port == "multi")], 1, 0)

ggplot(div_baci, aes(delta.simpson)) + geom_bar() + theme_classic()
ggplot(subset(div_baci, multi.port == 0), aes(delta.simpson)) + geom_bar() + theme_classic()

ggplot(subset(div_baci, multi.port == 0), aes(factor(ifq), delta.simpson)) + geom_boxplot()+ theme_classic()
ggplot(div_baci, aes(factor(ifq), delta.simpson)) + geom_boxplot()+ theme_classic()

ggplot(div_baci, aes(factor(multi.port), delta.simpson)) + geom_boxplot()+ theme_classic()

lm1 <- lm(delta.simpson ~ ifq, div_baci)
lm2 <- lm(delta.simpson ~ multi.port, div_baci)
lm3 <- lm(delta.simpson ~ multi.port + ifq, div_baci)
lm4 <- lm(delta.simpson ~ multi.port*ifq, div_baci)
```

## Community level impacts of management change

# Discussion


# Appendix
## Seasonal Fisheries
**FILL IN**

# References