---
title: "How many ports do vessels land at?"
author: "Emma Fuller"
date: "January 17, 2015"
output: html_document
---
```{r,echo=FALSE,warning=FALSE,message=FALSE}
tickets <- readRDS("/Users/efuller/1/CNH/Analysis/Metiers/results/2015-01-09/code/3_exploreBuildwebs/tickets.RDS")

library(reshape2)
max_port <- function(x,value.metric="revenue"){
  
  # function takes vessel ID and returns the port for which maximum revenue/lbs/trips were landed. 
  # value.metric = "revenue" "lbs" "trips"
  
  if(value.metric=="revenue"){
    melt_rev <- melt(x, id.vars = c("pcid","metier"), measure.vars = "adj_revenue")
  }else{
    if(value.metric=="lbs"){
      melt_rev <- melt(x, id.vars = c("pcid","metier"), measure.vars = "landed_wt")
  }else{ 
    if(value.metric=="trips"){
      melt_rev <- melt(x, id.vars = c("pcid","metier"), measure.vars = "trip_id")
  }else{
    warning("value.metric needs to be 'trip', 'revenue', or 'lbs'")
  }}}
  
  if(value.metric=="trips"){
    cast_rev <- dcast(melt_rev, pcid ~ metier,fun.aggregate = length)
  }else{
    cast_rev <- dcast(melt_rev, pcid ~ metier,fun.aggregate = sum)
  }
  
  row.names(cast_rev) <- cast_rev$pcid
  cast_rev$pcid <- NULL
  prop_rev <- rowSums(cast_rev)/sum(rowSums(cast_rev))
  maxes <- prop_rev[which.max(prop_rev)]
  
#   return(list(pcid = names(max_port), proportion =  max_port))
  return(maxes)
}

describe_portDist <- function(df){
  colnames(df)[2] <- "CRS"
  no_maj = length(which(df$CRS<.5))/nrow(df) # % vessels that don't have a majority port
  
  single_port = length(which(df$CRS==1))/nrow(df) # % vessels that have single port
  maj_port = length(which(df$CRS>.5 & df$CRS<1))/nrow(df) # % land more than 50% at one port, but still land at multiple ports
  return(list(no.majority = no_maj, single.port = single_port, major.port = maj_port))
}
```

Turns out that a fair number of vessels that really split among many ports

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(plyr)
prop_rev <- ddply(tickets, .(drvid), max_port)
lb_prop <- ddply(tickets, .(drvid), max_port, value.metric="lbs")
trip_prop <- ddply(tickets, .(drvid),max_port, value.metric="trips")
```

```{r,echo=FALSE,echo=FALSE,message=FALSE}
barplot(
  rbind(unlist(describe_portDist(prop_rev))*100, 
        unlist(describe_portDist(lb_prop))*100,
        unlist(describe_portDist(trip_prop))*100),
  beside = TRUE,col=c("lightseagreen","tomato","steelblue"),bor=F,
  ylab="percentage vessels",legend.text= c("revenue","lbs","trips"),args.legend=list(x="topleft",bty="n",bor=F))
```

+ `no.majority` means that a vessel does not land more than 50% of their trips (measured by revenue, lbs or number of trips) at any single port. 
+ `single.port` means that a vessel lands 100% of their trips (measured by revenue, lbs, or number of trips) at a single port
+ `major.port` means that a vessel lands between 50% and 100% of their trips at a single port. 

**Conclusion**: probably should drop vessels for which don't have at least 50% of trips in a single port, and come back to them later

### Details
Here I took each vessel and first found the percentage that it landed at each of the ports. Then I recorded the maximum of those percentages. I used that maximum percentage to determine whether it landed only at one port (maximum percentage = 1), had a majority (maximum percentage > .5 < 1) or had no majority (maximum percentage < .5). I did this for revenue, pounds landed and number of trips landed. 