---
title: "Metiers"
author: "Emma Fuller"
date: "May 6, 2014"
output: html_document
---

Just using FTL data (`grid` and `grgroup`) to do a clustering analysis

```{r,echo=FALSE, message=FALSE, warning=FALSE}

require(dplyr)
require(cluster)
load("/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-06/log_tripTable_2014-05-06.Rdata")
load("/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-04-28/new_plan/output_data/clust_log.Rdata")

# clustering solutions, if already exists don't re-run
file_find <- "/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-08/1_multi_gear/data/cluster_dat.Rdata"

if(file.exists(file_find)){
  load(file_find)
}else{ # generate clustering data
  effort_data <- log_tripTable[[2]]
  effort_df <- tbl_df(effort_data)
  effort_group <- group_by(effort_df, tripID)
  effort_codes<- summarise(effort_group,total_grgroup= paste(grgroup,collapse=""), total_grid = paste(grid,collapse=""))
  effort_codes$total_code <- paste(effort_codes$total_grgroup, effort_codes$total_grid,sep="_")
  effort_codes$total_code <- factor(effort_codes$total_code)
  
  # now check, same number of trips as trips with a unique gear code
  length(clust_log$clustering)==nrow(effort_codes)
  
  # make dataframe to cluster over
  cluster_dat <- data.frame(catch_profile = letters[clust_log$clustering], gear = effort_codes$total_code, tripID = effort_codes$tripID)
  
  #save input data
  save(cluster_dat, file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-08/1_multi_gear/data/cluster_dat.Rdata")
}

file_load <- "/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-08/1_multi_gear/data/clust_out.Rdata"
  
  
# cluster info
max.clusts <- 30
samples = 100
sampsize = 1000
clust.out <- vector("list",length=max.clusts)

set.seed(2)
for(i in 1:max.clusts){
  clust.out[[i]] <- clara(cluster_dat[,1:2], i, stand=TRUE, samples = samples, sampsize=sampsize, keep.data = FALSE, pamLike=TRUE, rngR = TRUE)
  print(i)
  }

save(clust.out, file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-08/1_multi_gear/data/clust_out.Rdata")

objectives <- vector(length = max.clusts)
for (i in 1:max.clusts) { objectives[i] <- clust.out[[i]]$objective }

asw <- vector(length = max.clusts)
for (i in 2:max.clusts) { asw[i] <- clust.out[[i]]$silinfo$avg.width  }
max_asw <- max(asw)
nodes <- which(asw==max(asw))
par(mfrow=c(1,2))
plot(objectives,type="o",pch=19,main="objective",xlab="Number of Clusters")
plot(asw,type="o",pch=19,main="asw",xlab="Number of Clusters")



plot(diff(asw),type='o')
points(22,asw[22],pch=19,col="red")

```
Before trying more nodes, what are these metiers made out of? 


```{r}
metier <- clust.out[[22]]$clustering

metier_sol <- cluster_dat
metier_sol$metier <- metier

barplot(table(metier_sol$gear,metier_sol$metier))
```

Looks like there are too many gear types. I need to drop out the multi-species gear. Looking at the gear types, there are gear types that account for 1-5 trips. Those rare ones are going to make it difficult to find clusters. 

```{r}
hist(table(metier_sol$gear)[table(metier_sol$gear)<1500],breaks=40,col="grey",bor="darkgrey",main="hist of num trips by gear",xlab="num trips")
```

## Remove multi-gear trips
First approach is to remove multi-gear trips

```{r}
effort_data <- log_tripTable[[2]]
effort_df <- tbl_df(effort_data)

# construct list of trips to remove
to_remove <- names(which(table(effort_group$tripID)>1))
# remove trips that match
remove_multi <- subset(effort_df,(!tripID %in% to_remove))
# remove miscellaneous trips
remove_multi <- subset(remove_multi,grgroup!="MSC")

# merge with catch profile
  # make catchprofile, tripID reference table
  ref <- log_tripTable[[2]][-which(duplicated(log_tripTable[[2]]$tripID)),]
  ref <- select(ref,tripID)
  ref$catch_profile <- clust_log$clustering

  # make gearID ref table
  gear_ref <- select(remove_multi,tripID, grgroup, grid)
  gear_ref$gear_code <- paste(gear_ref$grgroup,gear_ref$grid,sep="_")
  gear_ref$grgroup <- NULL
  gear_ref$grid <- NULL

  clust_dat <- merge(gear_ref, ref, by="tripID")
  clust_dat$catch_profile <- factor(letters[clust_dat$catch_profile])
  clust_dat$gear_code <- factor(clust_dat$gear_code)

# just cluster on grid and catch profile
if{file.exists(file_load)}{load(file_load)}else{
  # cluster info
max.clusts <- 30
samples = 100
sampsize = 1000
clust.out <- vector("list",length=max.clusts)

set.seed(2)
for(i in 1:max.clusts){
  clust.out[[i]] <- clara(clust_dat[,2:3], i, stand=TRUE, samples = samples, sampsize=sampsize, keep.data = FALSE, pamLike=TRUE, rngR = TRUE)
  print(i)
  }
 
objectives <- vector(length = max.clusts)
for (i in 1:max.clusts) { objectives[i] <- clust.out[[i]]$objective }

asw <- vector(length = max.clusts)
for (i in 2:max.clusts) { asw[i] <- clust.out[[i]]$silinfo$avg.width  }
max_asw <- max(asw)
nodes <- which(asw==max(asw))
plot(objectives,type="o")
plot(asw,type="o")
plot(diff(asw),type='o')
points(22,asw[22],pch=19,col="red")

metier_cluster <- list(objective = objectives, asw = asw, num_catchProfile = 24, sampsize=sampsize, samples=samples, cluster_sol = 22, clustering = clust.out[[22]])

save(metier_cluster,file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-08/metier_cluster.RData")
}
```

Looking at what are in these 22 metiers

```{r}
metier <-metier_cluster$clustering

metier_sol <- clust_dat
metier_sol$metier <- metier

gear_ref <- select(remove_multi,tripID, grgroup, grid)

metier_all <- merge(metier_sol, gear_ref, by="tripID")

paint <- colorRampPalette(brewer.pal(8,"Dark2"))(length(unique(metier_all$grgroup)))

barplot(table(metier_all$grgroup,metier_all$metier),col=paint,bor=paint,legend=TRUE,args.legend=list(bty="n",border=FALSE))

paint <- colorRampPalette(brewer.pal(8,"Dark2"))(length(unique(metier_all$catch_profile)))
paint <- colorRampPalette(c("red","blue"))(length(unique(metier_all$catch_profile)))
barplot(table(metier_all$catch_profile, metier_all$metier),col=paint, legend=TRUE,bor=paint,args.legend = list(ncol=4, bty="n",border=FALSE))
```

Adding in species abundances and gear types

```{r}
load("/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-06/lb_tripTable_2014-05-06.Rdata")

total_info <- merge(metier_all,lb_tripTable[[1]],by="tripID")

# for each metier, plot top 10 species and gear types
par(mfrow=c(1,2))
foo <- subset(total_info, metier==1)
species_foo <- foo[,7:ncol(foo)]
# find 0 rows, what's the distribution of species?
boxplot(species_foo[-which(colSums(species_foo)==0)])

# what's up with the sardine contingent?

sardines <- species_foo[which(species_foo$PSDN>0),]
sardines <- sardines[,-which(colSums(sardines)==0)]
boxplot(sardines)
# notice that there are no Dungeness crab here. So why do they show up? I think this is a problem in the catch profile. 

# look at the metier without sardines

no_sardines <- species_foo[-which(species_foo$PSDN>0),]
no_sardines <- no_sardines[,-which(colSums(no_sardines)==0)]
boxplot(no_sardines,ylim=c(0,10000),las=2)
# there's still a lot of stuff in here. What are the distributions of mean abundance

hist(colMeans(no_sardines),breaks=100000,xlim=c(0,1)) 
# lots of tiny means. 
barplot(colMeans(no_sardines)[-which(colMeans(no_sardines)<0.4)])

# look at NANC and SABL. Are they showing up in the same trip as DCRB? or it's also a mistake?

nanc_sabl <- species_foo[which(species_foo$NANC > 0 | species_foo$SABL > 0),]
boxplot(nanc_sabl[,-which(colSums(nanc_sabl)==0)],las=2)
barplot(colSums(nanc_sabl)[-which(colSums(nanc_sabl)==0)],las=2)
# unfortunately those catches are not zero. Let's look at the trips that catch both crab and sablefish/nanc

multi_crb <- subset(nanc_sabl, DCRB > 0)
# uh, turns out there is one freaking trip like this. 

# going back to original dataset. Find all trips that have both DCRB and SABL or DCRB and NANC

check_sabl <- subset(species_foo, DCRB > 0 & SABL >0) # there is one. 
check_nanc <- subset(species_foo, DCRB > 0 & NANC >0) # there are no rows. 

# re-visit that sablefish catch, what gear was it using?
multi <- subset(total_info, DCRB > 0 & SABL > 0)
# turns out there a bunch of trips that get both. let's look at which metier they tend to fall into

paint <- colorRampPalette(brewer.pal(8,"Dark2"))(length(unique(multi$catch_profile)))
barplot(table(multi$catch_profile,multi$metier),legend=TRUE,ylim=c(0,30),args.legend = list(bty="n",ncol=8),col=paint)
# this shows that dugeness crab + sablefish got put in a bunch of different catch profiles, which in turn were apportioned across different metiers. i.e. 1,2 and 9 all got one particular time, but 19, 20 and 21 got another. But regardless, this is not a common to have both crab and sablefish. 

# look at the gear
paint <- colorRampPalette(brewer.pal(8,"Dark2"))(length(unique(multi$gear_code)))

barplot(table(multi$gear_code, multi$metier)[-which(rowSums(table(multi$gear_code,multi$metier))==0),],col=paint,legend=TRUE,border=paint,args.legend = list(bty="n",border=FALSE,ncol=2))

# so this happened once using a crab pot. It happend frequently when trawling with a selective flat-fish trawl (FTS), sometimes with a midwater trawl (MDT) and occastionally with a roller trawl (RLT). 

# in conclusion, it seems like it's correct to label that one crab-pot occurance with the rest of the crab. ... Although it did catch a whole lot of sablefish
barplot(colSums(check_sabl[-which(check_sabl==0)]),bor=FALSE,col="indianred")

# looking back into FTL data to see what's going on with that trip
FTL <- read.csv("/Volumes/NOAA_Data/CNH/Data/Catch/FTL_2009-2013_2014-03-21.csv")

# find tripID
total_info$tripID[66175]

that_one_trip <- subset(FTL, VEID==1700000 & YEAR==2011 & MONTH==5 & DAY==31)
# he did catch a bunch of crab, along with some sablefish. Doesn't look like he's endorsed for sablefish though. 
```

So I feel okay with the one guy who caught sablefish in a crab pot, it was a real thing. And the rest of the occurances seem more based on the fact that they're trawling and will pick up some dungeness crab. Ok. But the midwater species: mackeral, sardines and anchovies, why are those there?


```{r}
# trying more nodes. 

more_max <- 60

clust.out <- vector("list",length=(more_max - max.clusts))

set.seed(2)
for(i in (max.clusts+1):more_max){
  clust.out[[i]] <- clara(cluster_dat, i, stand=TRUE, samples = samples, sampsize=sampsize, keep.data = FALSE, pamLike=TRUE, rngR = TRUE)
  print(i)
  }

objectives_more <- vector(length = max.clusts)
for (i in (max.clusts+1):more_max) { objectives_more[i] <- clust.out[[i]]$objective }

asw_more <- vector(length = max.clusts)
for (i in (max.clusts+1):more_max) { asw_more[i] <- clust.out[[i]]$silinfo$avg.width  }
max_asw_more <- max(asw_more)
nodes_more <- which(asw_more==max(asw_more))

total_obj <- c(objectives, objectives_more[(max.clusts+1):more_max])
total_asw <- c(asw, asw_more[(max.clusts+1):more_max])

metier_cluster <- list(objective = total_obj, asw = total_asw, num_catchProfile = 24, sampsize=sampsize, samples=samples)

save(metier_cluster,file="/Volumes/NOAA_Data/CNH/Analysis/Metiers/results/2014-05-08/metier_cluster.RData")

}
par(mfrow=c(1,2))
plot(metier_cluster$objective,type='o', pch=19, main="objective",bty="n",cex=0.5,xlab="Number of Clusters",ylab="Objective Function")
plot(metier_cluster$asw,type='o',main="asw",bty="n",pch=19,cex=0.5,xlab="Number of Clusters", ylab="Average Silhouette Width")
```

After clustering just on gear type from the fish ticket data, I don't see any convergence here. The average silhouette width just keeps going up, and the values are all really low. I think I probably need to add another predictor variable, or possibly drop those vessels that have multiple gear types. 
