---
title: "Defining Fisheries"
author: "Emma Fuller"
date: "September 17, 2014"
output:
#  ioslides_presentation:
#    highlight: espresso
#    fig_width: 20
#    fig_height: 15
#    self_contained: false
  slidy_presentation:
    highlight: espresso
    self_contained: false
    theme: journal
    fig.width: 10
    fig.height: 8
    css: mystyle.css
---

Introduction
====================
type: exclaim

Based on grey literature, fishermen interviews, discussions with NMFS scientists and managers I have a general impression of the structure of the fisheries on the west coast. The big points are 

+ There is a giant pacific whiting (hake) fishery -- single species, uses midwater trawls
+ There is a pink shrimp fishery -- very clean, double or single rigged shrimp trawls
+ There is a large crab fishery -- single species, uses crab pots
+ There is an albacore tuna fishery -- uses trolls, not observered, thus "clean"
+ There is a salmon fishery -- uses trools, not observed, thus "clean"
+ There is a pacific halibut fishery -- very small due to catch restrictions, "clean", uses troll
+ There is a fixed gear groundfish fishery 
  + Focuses on sablefish uses fish pots and/or long lines
+ There is a mobile gear groundfish fishery
  + Possibly nearshore and deeper water segments
  + Dover sole, sablefish, and thornyheads are a common target of deeper water
  
The more my revealed structure resembles what I know the better. The insight will come from how groundfish are split up. 


## Latent Dirichlet Analysis

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
library(noamtools)
library(ggplot2)
library(scales)
library(reshape2)
AIC_df <- readRDS(file="/Users/efuller/1/CNH/Analysis/Metiers/writing/presentations/Rutgers_9-2014/data/AIC_df.RDS")

plot(AIC_df$ks, -1*(AIC_df$AICs),type="o",pch=19, bty="n", xlab="Number of targets", ylab = "Relative parsimony of model (negative AIC)", col=alpha("steelblue", .75), lwd=3)

```

## Target composition

<iframe src="/Users/efuller/1/CNH/Analysis/Metiers/results/2014-09-06/topic_model_runs/astoria_lda_/index.html" width = 1250 height=750></iframe>

## Are there targets that are always targetted together?
Clustering targets by co-occurance

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
library(cluster)
library(vegan)
library(scales)
k = 1:22
dat <- readRDS("/Users/efuller/1/CNH/Analysis/Metiers/writing/presentations/Rutgers_9-2014/data/dat.RDS")
cluster_results <- list()

# transform data into hellinger

trip_trans <- decostand(dat$theta.hat, method = "hellinger")

for(i in 1:max(k)){
  cluster_results[[i]] <- clara(trip_trans, k[i], samples = 70, sampsize = 200)
}

avg_sil <- rep(NA, max(k))
for(i in 2:max(k)){
  avg_sil[i] <- cluster_results[[i]]$silinfo$avg.width
}

plot(avg_sil,type='o',cex=3,pch=19, xlab="number of clusters", ylab="average silhouette width", bty="n", col=alpha("steelblue",.75),lwd=3)
text(k, avg_sil, k, col="white", cex=.95,lwd=10)
```

## Cluster composition

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}

# 14 looks pretty good. look at mediods for each

cluster_sol <- cluster_results[[14]]$medoids
# how many trips for each
sol_trips <- table(cluster_results[[14]]$clustering)

# par(mfrow=c(3,5))
# for(i in 1:nrow(cluster_sol)){
#   barplot(cluster_sol[i,],bor=F, main = paste0(sol_trips[i]," trips\n", round((sol_trips[i]/length(cluster_results[[14]]$clustering)*100),2),"% of total trips"), cex.names=.75,ylim=c(0,1))
# }

df_sol <- melt(cluster_sol)
colnames(df_sol) <- c("cluster","target", "proportion")

paint <- rep(NA, 22)
paint[c(11,17, 18, 16,19,14,20,21,22,15)] <-"#2171b5"
paint[8] <- "#6baed6" # c mackeral
paint[15] <- "#3182bd" # j mackeral
paint[13] <-"#d95f02" # albacore
paint[4] <-"#e41a1c"  # crab
paint[10] <-"#e7298a" # pink shrimp
paint[2] <- "#7570b3" #ling and salmon
paint[1] <-"#8856a7" # bottom ground fish
paint[3] <- "#810f7c" # rockfish
paint[5] <- "#8c96c6" # whiting-ish
paint[12] <-"#8c96c6" #whiting, but also widow
paint[6] <- "#e6ab02" #species of concern (rockfish)
paint[7] <- "#7a0177" #sablefish 
paint[9] <- "#7570b3"  # pacific cod, some rockfish, ling code, canary, sole, skates, doesn't make a huge amount of sense
ggplot(df_sol, aes(x=target, y=proportion)) + geom_bar(stat="identity",fill=paint) + facet_wrap(~cluster)+ theme_bw() + theme(strip.background = element_blank()) 
```

## External validation: how does gear line up?

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=10, eval=FALSE}

cluster_gear <- readRDS("/Users/efuller/1/CNH/Analysis/Metiers/writing/presentations/Rutgers_9-2014/data/cluster_gear.RDS")
library(pander)
panderOptions("table.split.table", Inf)
pander(table(cluster_gear$gear, cluster_gear$cluster), style="rmarkdown")
```

# Define fisheries
Unique combinations of gear and fish target that happen at least 10 times. 

## Who's making which trips?

```{r, eval=FALSE}
library(igraph)

```

## Appendix: hellinger transformation

## Appendidix: clustering