---
title: "Masking VMS trajectories"
author: "Emma Fuller"
date: "November 2, 2014"
output: pdf_document
---

For most vessels, VMS remains on at all times, even when vessels are hauled out and driven to, and launched at, new ports. Thus it's not sufficient to filter for vessels that have inferred speed of zero as a check for "at port" vessels (since they might be merrily crusing down the highway at 60mph behind a ford pickup). Instead I use a high resolution dataset from NOAA[^1] to find any VMS points which fall inside of the coastline. 

Using San Francisco as an example, I plot the trajectory of one vessel to see the effects of buffering 3 nautical miles out. The result is as follows, where red points are on land, orange points are within the buffer, and grey points are at sea.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(sp); library(move)
load("../../../results/2014-10-29/2_coastline.Rdata")
load("../../../results/2014-10-29/2_bpoly.Rdata")
load('../../../results/2014-10-29/sfmov1.Rdata')

plot(sfmov1, type='o',cex=.15, lwd=.5, col="white")
plot(bpoly, col="steelblue",bor=F,add=T)
plot(WC, col="wheat",add=T,bor="white")
points(sfmov1, type='p',cex=.5, lwd=.5, col=land_paint,pch=19)
````

[^1]: cite plz

For this vessel, the buffer is obviously way too big. I think it's likely that I don't need a buffer at all, and instead just use the onland to mask points. Will do that. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
new_color <- ifelse(land_paint=="orange","grey",land_paint)
plot(sfmov1, type='o',cex=.15, lwd=.5, col="white")
plot(bpoly, col="steelblue",bor=F,add=T)
plot(WC, col="wheat",add=T,bor="white")
points(sfmov1, type='p',cex=.5, lwd=.5, col=new_color,pch=19)
```